### 2024-04-03 README.Rmd
init

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# criticalvalue

<!-- badges: start -->
<!-- badges: end -->

The goal of criticalvalue is to ...

## Installation

You can install the development version of criticalvalue like so:

``` r
# require(devtools)
devtools::install_github("filippogambarota/criticalvalue")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(criticalvalue)
# devtools::load_all()

# t-test

x <- rnorm(30, 0.5, 1)
y <- rnorm(30, 0, 1)

ttest <- t.test(x, y)
critical(ttest)

# cor.test
ctest <- cor.test(x, y)
critical(ctest)

# linear model (unstandardized)
z <- rnorm(30)
q <- rnorm(30)

dat <- data.frame(x, y, z, q)
fit <- lm(y ~ x + q + z, data = dat)
fit <- critical(fit)
summary(fit)

# linear model (standardized)
fit <- lm(y ~ x + q + z, data = dat)
fit <- critical(fit, standardize = TRUE)
summary(fit)
```

### 2024-04-03 R/criticalvalue-funs.R

@@ -1,0 +1,308 @@
+crit_from_t_t2s <- function(t = NULL, n1, n2, se = NULL,
+                            conf.level, 
+                            hypothesis,
+                            var.equal = FALSE){
+  if(!var.equal){
+    warning("When var.equal = FALSE the critical value calculated from t assume sd1 = sd2!")
+  }
+  alpha <- .get_alpha(conf.level, hypothesis)
+  df <- n1 + n2 - 2
+  tc <- abs(qt(alpha, df))
+  if(is.null(t)){
+    warning("When t is NULL, d cannot be computed, returning NA")
+    d <- NA
+  }else{
+    d <- t * (1/n1 + 1/n2)
+  }
+  dc <- tc * (1/n1 + 1/n2)
+  if(is.null(se)){
+    warning("When se = NULL bc cannot be computed, returning NA!")
+    bc <- NA
+  }else{
+    bc <- (tc * se)
+  }
+  out <- list(d = d, dc = dc, bc = bc)
+  return(out)
+}
+
+crit_from_data_t2s <- function(m1, m2, 
+                               sd1, sd2, 
+                               n1, n2, 
+                               conf.level,
+                               hypothesis,
+                               var.equal = FALSE){
+  alpha <- .get_alpha(conf.level, hypothesis)
+  b <- m1 - m2
+  if(!var.equal){ # welch
+    se1 <- sd1 / sqrt(n1)
+    se2 <- sd2 / sqrt(n2)
+    se <- sqrt((se1^2 + se2^2)/2)
+    s <- sqrt((sd1^2 * (n1 - 1) + sd2^2 * (n2 - 1))/(n1 + n2 - 2))
+    df <- se^4/(se1^4/(n1-1) + se2^4/(n2-1))
+  }else{ # standard
+    s <- sqrt((sd1^2 * (n1 - 1) + sd2^2 * (n2 - 1)) / (n1 + n2 - 2))
+    se <- s * (1/n1 + 1/n2)
+    df <- n1 + n2 - 2
+  }
+  tc <- abs(qt(alpha, df))
+  d <- b / s
+  bc <- tc * se
+  dc <- tc * (1/n1 + 1/n2)
+  out <- list(d = d, dc = dc, bc = bc, df = df)
+  return(out)
+}
+
+
+crit_from_t_t1s <- function(t = NULL, n, se = NULL,
+                            conf.level,
+                            hypothesis){
[…]

### 2024-04-03 R/criticalvalue-methods.R

@@ -1,0 +1,145 @@
+critical <- function(x, ...){
+  UseMethod("critical")
+}
+
+#' @export
+critical.htest <- function(x){
+  # all implemented subtype of htest objects
+  mtds <- c("Two Sample", "One Sample", "correlation", "Paired")
+  
+  if(all(!grepl(paste0(mtds, collapse = "|"), x$method))){
+    stop(
+      sprintf("method %s of class %s not implemented yet!",
+              x$method, class(x))
+    )
+  } else {
+    D <- insight::get_data(x)
+    conf.level <- attributes(x$conf.int)$conf.level
+    
+    hypothesis <- ifelse(x$alternative == "two.sided", "2t", "1t")
+    
+    alpha <- .get_alpha(conf.level, hypothesis)
+    df <- x$parameter
+    tc <- qt(alpha, df)
+    t <- x$statistic
+    se <- x$stderr
+    b <- x$estimate
+    method <- x$method
+  }
+  
+  if(grepl("correlation", method)){
+    n <- df + 2
+    cc <- critical_cor(n = n)
+    dc <- bc <- cc$rc
+    d <- b
+    class(x) <- c("critvalue", "ctest", "htest")
+  }else{ # if t-test
+    if(grepl("Two Sample", method)){
+      n <- tapply(D$x, D$y, length)
+      if(grepl("Welch", method)){
+        ss <- tapply(D$x, D$y, sd)
+        mm <- tapply(D$x, D$y, mean)
+        tt <- critical_t2s(m1 = mm[1], m2 = mm[2],
+                           sd1 = ss[1], sd2 = ss[2], 
+                           n1 = n[1], n2 = n[2], 
+                           var.equal = FALSE)
+      }else{
+        tt <- critical_t2s(t = t, se = se, n1 = n[1], n2 = n[2], var.equal = TRUE,
+                           conf.level = conf.level, hypothesis = hypothesis)
+      }
+    } else if(grepl("One Sample", method)){
+      n <- length(D$x)
+      tt <- critical_t1s(t = t, se = se, n = n, hypothesis = hypothesis, conf.level = conf.level)
+    } else if(grepl("Paired", method)){
+      n <- length(D$x)
+      r12 <- cor(D$x[D$y == 1], D$x[D$y == 2])
+      tt <- critical_t2sp(t = t, 
+                          se = se, 
[…]

### 2024-04-03 R/deprecated/criticalvalue-funs.R

@@ -1,0 +1,174 @@
+critical_ttest <- function(m1 = NULL, m2 = NULL,
+                           sd1 = NULL, sd2 = NULL,
+                           n1, n2,
+                           r12 = NULL,
+                           se = NULL,
+                           sed = NULL,
+                           t = NULL,
+                           type = "ts",
+                           var.equal = FALSE,
+                           denominator = "pooled",
+                           conf.level = 0.95,
+                           hypothesis = c("><", "<", ">")){
+  
+  # check arguments with multiple options
+  type <- match.arg(type, c("ts", "os", "ps"))
+  denominator <- match.arg(denominator, c("pooled", "differences"))
+  hypothesis <- match.arg(hypothesis)
+  
+  # get the alpha value for the critical value
+  alpha <- .get_alpha(conf.level, hypothesis)
+  
+  if(type == "ts"){ # two sample t test
+    # t, n1, n2 -> works but assume sd1 = sd2
+    df <- n1 + n2 - 2
+    if(is.null(t)){
+      if(is.null(m1) | is.null(m2) | is.null(sd1) | is.null(sd2)){
+        stop("When t is NULL m1, m2, sd1 and sd2 need to be provided!")
+      }
+      b <- m1 - m2
+      if(!var.equal){ # welch
+        se1 <- sd1 / sqrt(n1)
+        se2 <- sd2 / sqrt(n2)
+        se <- sqrt((se1^2 + se2^2)/2)
+        s <- sqrt((sd1^2 * (n1 - 1) + sd2^2 * (n2 - 1))/(n1 + n2 - 2))
+        df <- se^4/(se1^4/(n1-1) + se2^4/(n2-1))
+      }else{ # standard
+        s <- sqrt((sd1^2 * (n1 - 1) + sd2^2 * (n2 - 1)) / (n1 + n2 - 2))
+        se <- s * (1/n1 + 1/n2)
+      }
+    }else{
+      if(is.null(se)){
+        warning("bc cannot be calculated without se, returning NA")
+        s <- NA
+        se <- NA
+      }else{
+        s <- se / (1/n1 + 1/n2)
+      }
+    }
+  } else if(type == "os"){ # one sample
+    if(is.null(t)){
+      if(is.null(m1) | is.null(sd1)){
[…]

### 2024-04-03 R/deprecated/criticalvalue-methods.R

@@ -1,0 +1,142 @@
+critical <- function(x, ...){
+  UseMethod("critical")
+}
+
+#' @export
+critical.htest <- function(x, denominator = "pooled"){
+  denominator <- match.arg(denominator, c("pooled", "differences"))
+  # all implemented subtype of htest objects
+  mtds <- c("Two Sample", "One Sample", "correlation", "Paired")
+  
+  if(all(!grepl(paste0(mtds, collapse = "|"), x$method))){
+    stop(
+      sprintf("method %s of class %s not implemented yet!",
+              x$method, class(x))
+    )
+  } else{
+    D <- insight::get_data(x)
+    conf.level <- attributes(x$conf.int)$conf.level
+    
+    hypothesis <- switch(x$alternative,
+      "two.sided" = "><",
+      "greater" = ">",
+      "less" = "<"
+    )
+    
+    alpha <- .get_alpha(conf.level, hypothesis)
+    df <- x$parameter
+    tc <- qt(alpha, df)
+    t <- x$statistic
+    se <- x$stderr
+    b <- x$estimate
+    method <- x$method
+  }
+  
+  if(grepl("correlation", method)){
+    d <- b
+    dc <- bc <- tc / sqrt((df + tc^2))
+    class(x) <- c("critvalue", "ctest", "htest")
+  }else{ # if t-test
+    if(grepl("Two Sample", method)){
+      b <- b[1] - b[2] # x - y
+      n <- tapply(D$x, D$y, length)
+      if(grepl("Welch", method)){
+        ss <- tapply(D$x, D$y, sd)
+        s <- sqrt((ss[1]^2 + ss[2]^2)/2) # average sd
+      }else{
+        se_df <- sqrt(1/n[1] + 1/n[2]) # standard error without sd
+        s <- se / se_df
+      }
+    } else if(grepl("One Sample", method)){
+      n <- length(D$x)
+      se_df <- sqrt(1/n) # standard error without sd
+      s <- se / se_df
+    } else if(grepl("Paired", method)){
+      n <- tapply(D$x, D$y, length)
+      se_df <- sqrt(1/n[1]) # standard error without sd
+      ss <- tapply(D$x, D$y, sd)
+      if(denominator == "pooled"){
+        s <- sqrt((ss[1]^2 + ss[2]^2)/2)
+      } else if(denominator == "differences"){
+        s <- se / se_df
[…]

### 2024-04-03 R/extra.R

@@ -1,0 +1,72 @@
+zscore <- function(x, s = 1, center = TRUE, scale = TRUE, na.rm = FALSE){
+  m <- mean(x, na.rm = na.rm)
+  sd <- sd(x, na.rm = na.rm)
+  if(center){
+    x <- x - m
+  }
+  if(scale){
+    x <- x / (sd*s)
+  }
+  return(x)
+}
+
+.weighted.sd <- function(x, w){
+  n <- length(x)
+  sqrt(sum(((x - weighted.mean(x, w))*w)^2)/(n - 1))
+}
+
+standardize <- function(x, stand.y = TRUE, stand.x = TRUE){
+  
+  Y <- model.frame(x)[, 1, drop = FALSE]
+  X <- model.frame(x)[, -1, drop = FALSE]
+  
+  numIDX <- sapply(X, is.numeric)
+  
+  if(stand.y){
+    Y <- scale(Y)
+  }
+  if(stand.x){
+    for(i in 1:ncol(X)){
+      if(is.numeric(X[, i])){
+        X[, i] <- scale(X[, i])
+      }
+    }
+  }
+  D <- cbind(Y, X)
+  update(x, data = D)
+  
+}
+
+old_standardize <- function(x, 
+                        stand.x = TRUE, 
+                        stand.y = FALSE,
+                        nsd = 1){
+  Y <- model.frame(x)[, 1]
+  X <- model.matrix(x)
+  if(exists("weights", x)){
+    w <- x$weights
+  }else{
+    w <- rep(1, nrow(x$model))
+  }
+  
+  sY <- .weighted.sd(Y, w)
+  sX <- apply(X, 2, .weighted.sd, w)
+  
+  sY <- sY * nsd
+  sX <- sX * nsd
+  
+  B <- coef(x)
+  SE <- summary(x)$coefficients[, 2]
+  
+  if(stand.x){
+    B <- B * sX
+    SE <- SE * sX 
+  }
+  if(stand.y){
+    B <- B/sY
+    SE <- SE/sY
+  }
+  x$standardized.coefficients <- B
+  x$standardized.se <- SE
+  return(x)
+}

### 2024-04-03 R/utils.R

@@ -1,0 +1,44 @@
+.p_value <- function(q,
+                 df = NULL,
+                 test = c("t", "z"),
+                 conf.level = 0.95,
+                 hypothesis = c("1t", "2t")){
+  hypothesis <- match.arg(hypothesis)
+  test <- match.arg(test)
+  # two tails
+  if(test == "t"){
+    p <- 2 * pt(-abs(q), df)
+  }else{
+    p <- 2 * pnorm(-abs(q))
+  }
+  if(hypothesis == "1t"){
+    p <- p / 2
+  }
+  return(p)
+}
+
+.get_alpha <- function(conf.level = 0.95,
+                      hypothesis = c("2t", "1t")){
+  alpha <- 1 - conf.level
+  if(hypothesis == "2t"){
+    alpha <- alpha / 2
+  }
+  return(alpha)
+}
+
+.get_J <- function(df){
+  # taken from https://github.com/easystats/effectsize/blob/a579a5093549be2edc264e2d1da95726ff4f4369/R/cohens_d.R#L315
+  exp(lgamma(df/2) - log(sqrt(df/2)) - lgamma((df - 1)/2))
+}
+
+.round_list <- function(x, digits = getOption("digits")){
+  out <- lapply(x, function(el){
+    if(is.numeric(el)){
+      round(el, digits)
+    } else{
+      el
+    }
+  })
+  class(out) <- class(x)
+  out
+}

### 2024-04-03 deprecated/critical-value-class.R

@@ -1,0 +1,132 @@
+set.seed(2021)
+x <- rnorm(100)
+y <- rnorm(100)
+ht <- t.test(x, y)
+attributes(ht$conf.int)
+attr(ht, "class") <- c("feonyte", "htest")
+class(ht)
+
+print.feonyte <- function(x, ...){
+    NextMethod(x, ...)
+    alpha <- attr(x$conf.int, "conf.level")
+    cat("confidence level:", alpha, "\n\n")
+    invisible(NULL)
+}
+
+stats:::print.htest(ht)
+
+fit <- t.test(x, y, var.equal = TRUE)
+
+
+t.test()
+
+.all_null <- function(...){
+    all(sapply(list(...), is.null))
+}
+
+critical_t_test <- function(sd1 = NULL, 
+                            sd2 = NULL, 
+                            n1 = NULL,
+                            n2 = NULL,
+                            se = NULL, 
+                            df = NULL,
+                            alpha = 0.05,
+                            var.equal = FALSE){
+    if(is.null(se) & is.null(df)){
+        if(is.null(sd1) | is.null(sd2) | is.null(n1) | is.null(n2)){
+            stop("when se/df are NULL, sd1, sd2, n1 and n2 must be specified!")
+        }
+        stats <- t_test_se_df(n1, n2, sd1, sd2, var.equal)
+        se <- stats$se
+        df <- stats$df
+    }
+    tc <- critical_t(df, alpha/2)
+    critical_b <- se * tc
+    dc <- abs(critical_b)
+    list(critical = dc, se = se, df = df)
+}
+
+critical_cor_test <- function(n = NULL, alpha = 0.05){
+    df <- n - 2
+    tc <- critical_t(df, alpha/2)
+    abs(tc / sqrt(df + tc))
+}
+
+critical_t <- function(df, alpha = 0.05){
+    qt(alpha, df = df)
+}
+
+
+x <- rnorm(100)
+y <- rnorm(100)
+
+fit <- t.test(x, y)
+d <- critical_t_test(fit$stderr, fit$parameter)
+pt(d/fit$stderr, fit$parameter, lower.tail = FALSE)
+
+
+se <- sqrt(var(x)/100 + var(y)/100)
+fit$stderr
+
[…]

### 2024-04-03 deprecated/test.R

@@ -1,0 +1,20 @@
+t.test <- function(x, den = "pooled", ...){
+    res <- stats::t.test(x, ...)
+    alpha <- 1 - attributes(res$conf.int)$conf.level
+    nu <- res$parameter
+    res$tc <- qt(p = alpha/2, df = nu)
+    res$bc <- res$tc * res$stderr
+    class(res) <- c("critval", class(res))
+    return(res)
+}
+
+print.critval <- function(x, digits = getOption("digits"), prefix = "\t", ...){
+    NextMethod(x, ...)
+    critical_b <- sprintf("critical |b| = %.3f", abs(x$bc))
+    cat(critical_b, "\n")
+    invisible(NULL)
+}
+
+x <- t.test(rnorm(100), rnorm(100), paired = TRUE)
+
+x

### 2024-04-03 deprecated/test2.R

@@ -1,0 +1,30 @@
+get_htest <- function(x){
+    patterns <- c("ttest" = "t-test", 
+                  "cortest" = "correlation")
+    if(!"htest" %in% class(x)){
+        stop("x need to be of class htest!")
+    }else{
+        p <- sapply(patterns, function(p) grepl(p, x$method))
+        names(patterns)[p]
+    }
+}
+
+add_critical <- function(x){
+    htest <- get_htest(x)
+    class(x) <- c("critval", class(x))
+    return(x)
+}
+
+print.critval <- function(x, ...){
+    NextMethod(x, ...)
+    critical_b <- sprintf("critical b = %.3f", 0.45)
+    critical_d <- sprintf("critical d = %.3f", 0.1)
+    cat(critical_b)
+    invisible(NULL)
+}
+
+x <- t.test(rnorm(100))
+y <- cor.test(rnorm(100), rnorm(100))
+
+x <- add_critical(x)
+x

### 2024-04-03 example.R

@@ -1,0 +1,17 @@
+# devtools::load_all()
+# library(criticalvalue)
+
+# two sample t.test (var.equal = TRUE)
+
+x <- rnorm(100, 0.5, 1)
+y <- rnorm(100, 0, 1)
+
+tt <- t.test(x, y)
+fit <- lm(y ~ x)
+
+critical(fit, standardize = TRUE) |> 
+  summary()
+
+
+
+

### 2024-04-03 tests/testthat.R

@@ -1,0 +1,12 @@
+# This file is part of the standard setup for testthat.
+# It is recommended that you do not modify it.
+#
+# Where should you do additional test configuration?
+# Learn more about the roles of various files in:
+# * https://r-pkgs.org/testing-design.html#sec-tests-files-overview
+# * https://testthat.r-lib.org/articles/special-files.html
+
+library(testthat)
+library(criticalvalue)
+
+test_check("criticalvalue")

### 2024-04-03 tests/testthat/test-utils2.R

@@ -1,0 +1,151 @@
+# cor test, t, conf.level = 0.95 -------------------------------
+
+test_that("critical value for correlation works (t, two sided)", {
+  n <- 40
+  rr <- critical_cortest(n, test = "t")
+  t <- rr$rc / rr$se_rc
+  # should be 0.05
+  pval <- p_value(t, rr$df, test = "t", hypothesis = "><")
+  expect_equal(pval, 0.05)
+})
+
+test_that("critical value for correlation works (t, greater)", {
+  n <- 40
+  rr <- critical_cortest(n, test = "t", hypothesis = ">")
+  t <- rr$rc / rr$se_rc
+  # should be 0.05
+  pval <- p_value(t, rr$df, test = "t", hypothesis = ">")
+  expect_equal(pval, 0.05)
+})
+
+test_that("critical value for correlation works (t, less)", {
+  n <- 40
+  rr <- critical_cortest(n, test = "t", hypothesis = "<")
+  t <- rr$rc / rr$se_rc
+  # should be 0.05
+  pval <- p_value(t, rr$df, test = "t", hypothesis = "<")
+  expect_equal(pval, 0.05)
+})
+
+# cor test, t, conf.level = 0.80 -------------------------------
+
+test_that("critical value for correlation works (t, two sided)", {
+  n <- 40
+  conf.level <- 0.8
+  rr <- critical_cortest(n, test = "t", conf.level = conf.level)
+  t <- rr$rc / rr$se_rc
+  pval <- p_value(t, rr$df, test = "t", hypothesis = "><", conf.level = conf.level)
+  pval_expected <- 1 - conf.level
+  expect_equal(pval, pval_expected)
+})
+
+test_that("critical value for correlation works (t, greater)", {
+  n <- 40
+  conf.level <- 0.8
+  rr <- critical_cortest(n, test = "t", hypothesis = ">", conf.level = conf.level)
+  t <- rr$rc / rr$se_rc
+  pval_expected <- 1 - conf.level
+  pval <- p_value(t, rr$df, test = "t", hypothesis = ">", conf.level = conf.level)
+  expect_equal(pval, pval_expected)
+})
+
+test_that("critical value for correlation works (t, less)", {
+  n <- 40
+  conf.level <- 0.8
[…]

### 2024-04-03 README.Rmd
fixing readme

@@
 ```{r example}
-library(criticalvalue)
-# devtools::load_all()
+# library(criticalvalue)
+devtools::load_all()
 
@@
 fit <- critical(fit)
+fit
 summary(fit)

### 2024-04-03 R/criticalvalue-methods.R
fixing missing gzc and gz

@@ -41,7 +41,8 @@ critical.htest <- function(x){
         mm <- tapply(D$x, D$y, mean)
         tt <- critical_t2s(m1 = mm[1], m2 = mm[2],
                            sd1 = ss[1], sd2 = ss[2], 
-                           n1 = n[1], n2 = n[2], 
+                           n1 = n[1], n2 = n[2],
+                           hypothesis = hypothesis,
                            var.equal = FALSE)
       }else{
         tt <- critical_t2s(t = t, se = se, n1 = n[1], n2 = n[2], var.equal = TRUE,
@@ -61,6 +62,8 @@ critical.htest <- function(x){
                           conf.level = conf.level)
       x$dz <- tt$dz
       x$dzc <- tt$dzc
+      x$gz <- tt$gz
+      x$gzc <- tt$gzc
     }
 
     d <- tt$d
@@ -112,8 +115,15 @@ print.critvalue <- function(x, digits = getOption("digits"), ...){
   x <- .round_list(x, digits)
   if(inherits(x, "ttest")){
     cat("|== Effect Size and Critical Value ==|", "\n")
-    cat("d =", x$d, "|dc| =", abs(x$dc), "|bc| =", abs(x$bc),"\n")
-    cat("g =", x$g, "|gc| =", abs(x$gc), "\n\n")
+    if(grepl("Paired", x$method)){
+      cat("d =", x$d, "|dc| =", abs(x$dc), "|bc| =", abs(x$bc),"\n")
+      cat("dz =", x$dz, "|dzc| =", abs(x$dzc), "\n")
+      cat("gz =", x$gz, "|gzc| =", abs(x$gzc), "\n\n")
+    }else{
+      cat("d =", x$d, "|dc| =", abs(x$dc), "|bc| =", abs(x$bc),"\n")
+      cat("g =", x$g, "|gc| =", abs(x$gc), "\n\n")
+    }
+    
   }else if(inherits(x, "ctest")){
     cat("|== Critical Value ==|", "\n")
     cat("|rc| =", abs(x$dc),"\n\n")

### 2024-04-03 R/criticalvalue-funs.R
adding greater and less not only 1t

@@ -159,7 +159,7 @@ crit_from_data_t2sp <- function(m1, m2 = NULL,
 
 critical_t1s <- function(m = NULL, s = NULL, t = NULL,
                          n, se = NULL,
-                         hypothesis = c("2t", "1t"),
+                         hypothesis = c("two.sided", "greater", "less"),
                          conf.level = 0.95){
   hypothesis <- match.arg(hypothesis)
   if(!is.null(m)){
@@ -185,7 +185,7 @@ critical_t2s <- function(m1 = NULL, m2 = NULL, t = NULL,
                          sd1 = NULL, sd2 = NULL,
                          n1, n2, se = NULL,
                          var.equal = FALSE,
-                         hypothesis = c("2t", "1t"),
+                         hypothesis = c("two.sided", "greater", "less"),
                          conf.level = 0.95){
   hypothesis <- match.arg(hypothesis)
   if(!is.null(m1) | !is.null(m2)){
@@ -219,7 +219,7 @@ critical_t2s <- function(m1 = NULL, m2 = NULL, t = NULL,
 critical_t2sp <- function(m1 = NULL, m2 = NULL, t = NULL,
                           sd1 = NULL, sd2 = NULL, r12 = NULL,
                           n, se = NULL,
-                          hypothesis = c("2t", "1t"),
+                          hypothesis = c("two.sided", "greater", "less"),
                           conf.level = 0.95){
   hypothesis <- match.arg(hypothesis)
   if(!is.null(m1)){
@@ -247,7 +247,7 @@ critical_t2sp <- function(m1 = NULL, m2 = NULL, t = NULL,
 
 critical_cor <- function(r = NULL, n, 
                          conf.level = 0.95, 
-                         hypothesis = c("2t", "1t"), 
+                         hypothesis = c("two.sided", "greater", "less"), 
                          test = c("t", "z")){
   df <- n - 2
   test <- match.arg(test, test)
@@ -280,7 +280,7 @@ critical_cor <- function(r = NULL, n,
 
 critical_coef_lm <- function(seb, n = NULL, p = NULL,df = NULL,
                              conf.level = 0.95,
-                             hypothesis = c("2t", "1t"),
+                             hypothesis = c("two.sided", "greater", "less"),
                              test = c("t", "z")){
   if(is.null(df) & (is.null(p) & is.null(n))){
     stop("df or p and n need cannot be NULL")

### 2024-04-03 R/criticalvalue-methods.R

@@ -16,7 +16,8 @@ critical.htest <- function(x){
     D <- insight::get_data(x)
     conf.level <- attributes(x$conf.int)$conf.level
     
-    hypothesis <- ifelse(x$alternative == "two.sided", "2t", "1t")
+    # hypothesis <- ifelse(x$alternative == "two.sided", "2t", "1t")
+    hypothesis <- x$alternative
     
     alpha <- .get_alpha(conf.level, hypothesis)
     df <- x$parameter
@@ -29,7 +30,7 @@ critical.htest <- function(x){
   
   if(grepl("correlation", method)){
     n <- df + 2
-    cc <- critical_cor(n = n)
+    cc <- critical_cor(n = n, conf.level = conf.level, hypothesis = hypothesis)
     dc <- bc <- cc$rc
     d <- b
     class(x) <- c("critvalue", "ctest", "htest")
@@ -83,12 +84,12 @@ critical.htest <- function(x){
   return(x)
 }
 #' @export
-critical.lm <- function(x, conf.level = 0.95, 
-                        hypothesis = c("2t", "1t"),
+critical.lm <- function(x, conf.level = 0.95,
                         standardize = FALSE,
                         ...){
   
-  hypothesis <- match.arg(hypothesis)
+  # always two.sided
+  hypothesis <- "two.sided"
   alpha <- .get_alpha(conf.level, hypothesis)
   df <- x$df.residual
   
@@ -115,18 +116,40 @@ print.critvalue <- function(x, digits = getOption("digits"), ...){
   x <- .round_list(x, digits)
   if(inherits(x, "ttest")){
     cat("|== Effect Size and Critical Value ==|", "\n")
-    if(grepl("Paired", x$method)){
+    if(x$alternative == "two.sided"){
       cat("d =", x$d, "|dc| =", abs(x$dc), "|bc| =", abs(x$bc),"\n")
-      cat("dz =", x$dz, "|dzc| =", abs(x$dzc), "\n")
-      cat("gz =", x$gz, "|gzc| =", abs(x$gzc), "\n\n")
+      cat("g =", x$g, "|gc| =", abs(x$gc), "\n")
+    } else if(x$alternative == "greater"){
+      cat("d =", x$d, "dc =", abs(x$dc), "bc =", abs(x$bc),"\n")
[…]

### 2024-04-03 R/utils.R

@@ -19,6 +19,7 @@
 
 .get_alpha <- function(conf.level = 0.95,
                       hypothesis = c("2t", "1t")){
+  hypothesis <- if(hypothesis == "two.sided") "2t" else "1t"
   alpha <- 1 - conf.level
   if(hypothesis == "2t"){
     alpha <- alpha / 2
@@ -41,4 +42,9 @@
   })
   class(out) <- class(x)
   out
+}
+
+.hyp_sign <- function(x, hypothesis){
+  x <- abs(x)
+  ifelse(hypothesis %in% c("two.sided", "greater"), x, -x) 
 }

### 2024-04-04 R/criticalvalue-funs.R
fixing d from t equation and passing se from object

@@ -12,16 +12,16 @@ crit_from_t_t2s <- function(t = NULL, n1, n2, se = NULL,
     warning("When t is NULL, d cannot be computed, returning NA")
     d <- NA
   }else{
-    d <- t * (1/n1 + 1/n2)
+    d <- t * sqrt(1/n1 + 1/n2)
   }
-  dc <- tc * (1/n1 + 1/n2)
+  dc <- tc * sqrt(1/n1 + 1/n2)
   if(is.null(se)){
     warning("When se = NULL bc cannot be computed, returning NA!")
     bc <- NA
   }else{
     bc <- (tc * se)
   }
-  out <- list(d = d, dc = dc, bc = bc)
+  out <- list(d = d, dc = dc, bc = bc, df = df)
   return(out)
 }
 
@@ -30,24 +30,28 @@ crit_from_data_t2s <- function(m1, m2,
                                n1, n2, 
                                conf.level,
                                hypothesis,
+                               se = NULL,
+                               df = NULL,
                                var.equal = FALSE){
   alpha <- .get_alpha(conf.level, hypothesis)
   b <- m1 - m2
   if(!var.equal){ # welch
     se1 <- sd1 / sqrt(n1)
     se2 <- sd2 / sqrt(n2)
-    se <- sqrt((se1^2 + se2^2)/2)
-    s <- sqrt((sd1^2 * (n1 - 1) + sd2^2 * (n2 - 1))/(n1 + n2 - 2))
-    df <- se^4/(se1^4/(n1-1) + se2^4/(n2-1))
+    if(is.null(se)) se <- sqrt(se1^2 + se2^2)
+    # average sd
+    s <- sqrt((sd1^2 + sd2^2)/2)
+    if(is.null(df)) df <- se^4/(se1^4/(n1-1) + se2^4/(n2-1))
   }else{ # standard
+    # pooled sd
     s <- sqrt((sd1^2 * (n1 - 1) + sd2^2 * (n2 - 1)) / (n1 + n2 - 2))
-    se <- s * (1/n1 + 1/n2)
-    df <- n1 + n2 - 2
+    if(is.null(se)) se <- s * (1/n1 + 1/n2)
+    if(is.null(df)) df <- n1 + n2 - 2
   }
   tc <- abs(qt(alpha, df))
   d <- b / s
   bc <- tc * se
-  dc <- tc * (1/n1 + 1/n2)
+  dc <- tc * sqrt(1/n1 + 1/n2)
[…]

### 2024-04-04 R/criticalvalue-methods.R
removing space in print.htest

@@ -43,6 +43,7 @@ critical.htest <- function(x){
         tt <- critical_t2s(m1 = mm[1], m2 = mm[2],
                            sd1 = ss[1], sd2 = ss[2], 
                            n1 = n[1], n2 = n[2],
+                           se = se,
                            hypothesis = hypothesis,
                            var.equal = FALSE)
       }else{
@@ -121,10 +122,10 @@ print.critvalue <- function(x, digits = getOption("digits"), ...){
       cat("g =", x$g, "|gc| =", abs(x$gc), "\n")
     } else if(x$alternative == "greater"){
       cat("d =", x$d, "dc =", abs(x$dc), "bc =", abs(x$bc),"\n")
-      cat("g =", x$g, "gc =", abs(x$gc), "\n\n")
+      cat("g =", x$g, "gc =", abs(x$gc), "\n")
     }else{
       cat("d =", x$d, "dc =", -abs(x$dc), "bc =", -abs(x$bc),"\n")
-      cat("g =", x$g, "gc =", -abs(x$gc), "\n\n")
+      cat("g =", x$g, "gc =", -abs(x$gc), "\n")
     }
     
     if(grepl("Paired", x$method)){

### 2024-04-04 README.Rmd
updating readme
updating readme with examples

@@
 )
+
+options(scipen = 999, digits = 5)
 ```
@@
 
-## Example
+## Installation (temporary)
+
+The package is not production-ready thus to test the feature we suggest to clone or download the repository, open the R Project and load the functions with `devtools::load_all()`.
+
+## Example from fitted objects
 
@@
 
+### T Test
+
 ```{r example}
-# library(criticalvalue)
-devtools::load_all()
+# library(criticalvalue) # when the package will be ready
+devtools::load_all() # loading all functions for testing
 
-# t-test
+# t-test (welch)
 
@@
 
+# t-test (standard)
+
+ttest <- t.test(x, y, var.equal = TRUE)
+critical(ttest)
+
+# t-test (standard) with monodirectional hyp
+
+ttest <- t.test(x, y, var.equal = TRUE, alternative = "less")
+critical(ttest)
+
+# within the t-test object saved from critical we have all the new values
+
+ttest <- critical(ttest)
+str(ttest)
+```
+
+We can check the results using:
+
+```{r}
+ttest <- t.test(x, y)
+ttest <- critical(ttest)
+
+t <- ttest$bc/ttest$stderr # critical numerator / standard error
+
+# should be 0.05 (or alpha)
+(1 - pt(ttest$bc/ttest$stderr, ttest$parameter)) * 2
+```
+
+### Correlation Test
+
+```{r}
 # cor.test
@@
 critical(ctest)
+```
 
+### Linear Model
+
+```{r}
 # linear model (unstandardized)
@@
 fit <- critical(fit, standardize = TRUE)
+fit
 summary(fit)
 ```
+
+## Example from summary statistics
+
+```{r}
+x <- rnorm(30, 0.5, 1)
+y <- rnorm(30, 0, 1)
+
+ttest <- t.test(x, y)
+
+m1 <- mean(x)
+m2 <- mean(y)
+sd1 <- sd(x)
+sd2 <- sd(y)
+n1 <- n2 <- 30
+
+critical_t2s(m1, m2, sd1 = sd1, sd2 = sd2, n1 = n1, n2 = n2)
+
+critical_t2s(t = ttest$statistic, se = ttest$stderr, n1 = n1, n2 = n2)
+
+critical_t2s(t = ttest$statistic, n1 = n1, n2 = n2)
+```

### 2024-04-07 R/criticalvalue-funs.R
critical_coef_lm to critical_coef for rma

@@ -287,7 +287,7 @@ critical_cor <- function(r = NULL, n,
   return(out)
 }
 
-critical_coef_lm <- function(seb, n = NULL, p = NULL,df = NULL,
+critical_coef <- function(seb, n = NULL, p = NULL,df = NULL,
                              conf.level = 0.95,
                              hypothesis = c("two.sided", "greater", "less"),
                              test = c("t", "z")){
@@ -313,5 +313,3 @@ critical_coef_lm <- function(seb, n = NULL, p = NULL,df = NULL,
   out <- list(bc = bc, test = test)
   return(out)
 }
-
-

### 2024-04-07 R/criticalvalue-methods.R
adding critical.rma

@@ -84,6 +84,7 @@ critical.htest <- function(x){
   
   return(x)
 }
+
 #' @export
 critical.lm <- function(x, conf.level = 0.95,
                         standardize = FALSE,
@@ -99,13 +100,32 @@ critical.lm <- function(x, conf.level = 0.95,
   }
   
   se <- sqrt(diag(vcov(x)))
-  ll <- critical_coef_lm(se, df = df, conf.level = conf.level, hypothesis = hypothesis)
+  ll <- critical_coef(se, df = df, conf.level = conf.level, hypothesis = hypothesis)
   d <- NA
   dc <- NA
   x$d <- NA
   x$dc <- NA
   x$bc <- unname(ll$bc)
   class(x) <- c("critvalue", "lm")
+  return(x)
+}
+
+#' @export
+critical.rma <- function(x, conf.level = 0.95, h0 = 0){
+  if(inherits(x, "rma.uni")){
+    hypothesis <- "two.sided"
+    se <- x$se
+    df <- x$k.eff
+    ll <- critical_coef(se, df = df, conf.level = conf.level, hypothesis = hypothesis)
+    d <- NA
+    dc <- NA
+    x$d <- NA
+    x$dc <- NA
+    x$bc <- unname(ll$bc)
+    class(x) <- c("critvalue", "rma.uni", "rma")
+  } else{
+    stop(paste("class", class(x)[1], "not supported yet!"))
+  }
   return(x)
 }
 
@@ -157,6 +177,14 @@ print.critvalue <- function(x, digits = getOption("digits"), ...){
     names(bc) <- names(coef(x))
     print(abs(bc), ...)
     cat("\n")
+  } else if(inherits(x, "rma.uni")){
+    crit <- data.frame(x$b)
+    crit[, 1] <- x$bc
+    if(x$int.only){
+      rownames(crit) <- ""
+    }
+    names(crit) <- "|critical estimate|"
+    cat(capture.output(print(crit)), sep = "\n")
   }
   invisible(x)
 }
@@ -164,16 +192,17 @@ print.critvalue <- function(x, digits = getOption("digits"), ...){
 #' @export
 summary.critvalue <- function(x, ...){
   NextMethod(x, ...)
-  # taken from https://github.com/cran/lm.beta/blob/master/R/summary.lm.beta.R
-  x2 <- x
-  attr(x2, "class") <- "lm"
-  x.summary <- summary(x2, ...)
-  x.summary$coefficients <- cbind(x.summary$coefficients[, 
[…]

### 2024-06-03 example.R
Update

@@ -36,5 +36,6 @@ pwr::pwr.t.test(n=500,power=0.8)
 metafor:::.cmicalc()
 
 
+# prova di modifica
 
-
+#perchè non funziona

### 2024-06-03 Examples4paper.R
Provo a caricare il nuovo file

@@ -1,0 +1,16 @@
+devtools::load_all()
+
+# require(devtools)
+devtools::install_github("filippogambarota/criticalvalue")
+
+library(psych)
+library(psychTools)
+
+data("holzinger.swineford")
+Holz <- holzinger.swineford
+
+tt <- t.test(Holz$t01_visperc [ Holz$female == 1], Holz$t01_visperc [ Holz$female == 2])
+critical.htest(tt)
+
+
+

### 2024-06-03 example.R

@@ -36,6 +36,3 @@ pwr::pwr.t.test(n=500,power=0.8)
 metafor:::.cmicalc()
 
 
-# prova di modifica
-
-#perchè non funziona

### 2024-06-03 Examples4paper.R
New updates

@@ -9,8 +9,50 @@ library(psychTools)
 data("holzinger.swineford")
 Holz <- holzinger.swineford
 
+## Correlation
+
+cc<-cor.test(Holz$t01_visperc, Holz$t02_cubes)
+critical(cc)
+critical.htest(cc)
+
+## T-test
+
 tt <- t.test(Holz$t01_visperc [ Holz$female == 1], Holz$t01_visperc [ Holz$female == 2])
+critical(tt)
 critical.htest(tt)
+
+## Linear regression
+
+ll <- lm( t01_visperc ~ ageyr , data = Holz)
+summary(critical(ll))
+
+## T-test da n
+
+n <- 20
+
+critical_t2s(n1 = n , n2 = n , hypothesis = "less")
+
+## Paired t-test
+
+critical_t2sp(n = n , hypothesis = "less")
+
+## Critical correlation
+
+critical_cor(n = n,hypothesis = "less", test = "z")
+
+
+## Critical Coefficient
+
+critical_coef(seb = .1, n = 20, p = .01, conf.level = .95, hypothesis = "less")
+
+
+
+
+
+
+
+
+
 
 
 

### 2024-06-04 Examples.Rnw
new examples

\documentclass{article}

\begin{document}

<<>>=
devtools::load_all()

# require(devtools)
devtools::install_github("filippogambarota/criticalvalue")

@

<<>>=
library(psych)
library(psychTools)

data("holzinger.swineford")
Holz <- holzinger.swineford
@

Critical value for correlation from real data

<<>>=
cc<-cor.test(Holz$t01_visperc, Holz$t02_cubes)
critical(cc)
critical.htest(cc)
@

Critical value for t-test from real data

<<>>=
tt <- t.test(Holz$t01_visperc [ Holz$female == 1], Holz$t01_visperc [ Holz$female == 2])
critical(tt)
critical.htest(tt)
@

Critical value for linear model from real data

<<>>=
ll <- lm( t01_visperc ~ ageyr , data = Holz)
summary(critical(ll))
@

Critical t value from sample size

<<>>=
n <- 20
critical_t2s(n1 = n , n2 = n , hypothesis = "less")
@

Critical t value from sample size for a paired t-test

<<>>=
critical_t2sp(n = n , hypothesis = "less")
@

Critical value for a correlation from sample size

<<>>=
critical_cor(n = n,hypothesis = "less", test = "z")
@

Critical value of a beta from a model

<<>>=
critical_coef(seb = .1, n = 20, p = .01, conf.level = .95, hypothesis = "less")
@

\end{document}

### 2024-06-04 Examples.Rnw
update

@@
 \begin{document}
+
+To install the package the following line of code is needed.
 
@@
 
+Four our examples on real data we used prom the package 'psych' the dataset "holzinger.swineford" which has a series of demografics and scores of different subtests measuring intelligence.
+Once the package is retrieved with the 'library' command, the dataset can be opened using the 'data' function.
+For simplicity we decided to rename it with a shorter name.
+
 <<>>=
@@
 
-Critical value for correlation from real data
+Critical value for correlation from a cor.test object.
 
@@
 @
+
+In this first example we show how to calculate the critical value for a correlation test.
+We correlated two cognitive measures: visual perception and the cubes sub-test.
+It can be done by using firstly the 'cor.test' function and then putting the newly created object into the 'critical' function.
+The output will add, underneath the usual output of the 'cor.test' function, the absolute critical value (|rc|).
+The number represents the smallest significant effect.
+The same output can be obtained with the function 'critical.htest'.
 
@@
 
+In this example we performed a t-test comparing boys and girls on a cognitive variable of visual perception, as we assume gender differences.
+In this case, it can be easily done with the same procedure using the 't.test' function and putting the object inside either 'critical' or 'critical.htest'.
+The output now gives a wider range of values: the cohen's d calculated on the data (d), the critical cohen's d (dc), the numerator of the formula for the critical cohen's d (bc), the cohen's d adjusted for small samples (g) and the critical cohen's d adjusted for small samples (gc).
+
 Critical value for linear model from real data
@@
 
+In this last example, we expect the cognitive task to vary as a function of age.
+As the output shows, there is no relation between the sub-test and age because the scores are standardized.
+First, the model needs to be created and saved in an object, then it should be passed into the 'critical' function and wrapped in the 'summary' function.
+
+In the next examples we will show the use of the package's functions to calculate critical values in a prospective framework.
+Imagine having to plan a study and there can be different scenarios.
+Given external constrains on sample size, the question could be on which model can be implemented to find a significant interest, meaning that it might be the case that with a small sample, it would be better to opt for a simples study design.
+Or in another scenario we want to know the sample that would allow us to get a significant result under certain conditions and the smallest significant effect could tell us if the planned sample gives a realistic smallest effect.
+With a sample of 30 subjects, the critical cohen's d is of .62, which might be unrealistic for certain psychological variables.
+In an opposite scenario, with a sample size of 2400 subjects, the critical d would be of .06, would that be relevant in a comparison of gender differences?
+It is important to think about what sample we need to find a significant effect given a reasonable effect of interest.
+In the following examples we will give an outlook of the use of the functions to answer the aforesaid questions.
+
 Critical t value from sample size
@@
 <<>>=
-n <- 20
+n <- 1200
 critical_t2s(n1 = n , n2 = n , hypothesis = "less")
 @
+
+This example shows how to compute a critical value for a t-test with the 'critical_t2s' function in which the hypothesis can be specified as "less", "greater" and "two.tailed".
+In output we can find the critical d and the critical g, the adjusted d for small samples.
 
@@
 <<>>=
+n <- 15
 critical_t2sp(n = n , hypothesis = "less")
 @
+
+In the case of test-retest, for example the pre-post treatment, we are comparing the same subjects at two different time points.
+It is possible with the 'critical_t2sp' function to calculate the standardized and un-standardized critical cohen's d, and the critical cohen's d adjusted for small samples both standardized and un-standardized.
 
@@
 <<>>=
-critical_cor(n = n,hypothesis = "less", test = "z")
+n <- 60
+critical_cor(n = n, hypothesis = "two.sided", test = "z")
 @
+
+For a correlation, the function 'critical_cor' can be used to get the critical value specifying the sample size, the direction of the hypothesis and the test to apply, either t-test or z-test.
 
@@
 <<>>=
-critical_coef(seb = .1, n = 20, p = .01, conf.level = .95, hypothesis = "less")
+n <- 170
+critical_coef(seb = .1, n = n, p = .01, conf.level = .95, hypothesis = "less")
 @
 
+With the function 'critical_coef' one can calculate the critical beta by specifying sample size, standard error of the estimate, p-value, confidence level and hypothesis.
+
 \end{document}

### 2024-06-04 R/criticalvalue-funs.R

@@ -219,7 +219,7 @@ crit_from_t_t2sp <- function(t = NULL, n, se = NULL, r12 = NULL, hypothesis, con
 #' @param conf.level confidence level of the interval.
 #' @param hypothesis a character string indicating the alternative hypothesis ("less", "greater" or "two.tailed").
 #'
-#' @return the output returns a "dz" which is the d standartized on the standard deviation of the differencens, the "dzc" is the critical standardized d, the "d" is the cohen's d, the critical d which is the minimum value for which to get a significant result with a given sample, the "bc" is the numerator of the formula from which the d is calculated and "df" are the degrees of freedom.
+#' @return the output returns a "dz" which is the d standartized on the standard deviation of the differences, the "dzc" is the critical standardized d, the "d" is the cohen's d, the critical d which is the minimum value for which to get a significant result with a given sample, the "bc" is the numerator of the formula from which the d is calculated and "df" are the degrees of freedom.
 #' @export
 #'
 #' @examples
@@ -376,7 +376,7 @@ critical_t2s <- function(m1 = NULL, m2 = NULL, t = NULL,
 #' @param hypothesis a character string indicating the alternative hypothesis ("less", "greater" or "two.tailed").
 #' @param conf.level the confidence level to set the confidence interval, default is set to 0.95.
 #'
-#' @return the output returns, based on the inserted values, means of the two groups, standard deviations of the two groups, numerosity, standard error, correlation between the variables, confidence level and hypothesis or t-value, numerosity, standard error, correlation between the variables, hypothesis and confidence iterval.
+#' @return the output returns, based on the inserted values, means of the two groups, standard deviations of the two groups, numerosity, standard error, correlation between the variables, confidence level and hypothesis or t-value, numerosity, standard error, correlation between the variables, hypothesis and confidence interval.
 #' @export
 #'
 #' @examples

### 2024-06-04 ExPro.Rmd
new file

--- 2024-06-04 Examples.Rnw
@@
-\documentclass{article}
+---
+title: "Examples for Paper"
+output: html_document
+---
 
-\begin{document}
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
 
@@
 
-<<>>=
+```{r load}
 devtools::load_all()
@@
 # require(devtools)
-devtools::install_github("filippogambarota/criticalvalue")
+#devtools::install_github("filippogambarota/criticalvalue")
 
-@
+```
+
+## Examples on real data
 
@@
 
-<<>>=
+```{r open data}
 library(psych)
@@
 Holz <- holzinger.swineford
-@
+```
 
@@
 
-<<>>=
+```{r correlation}
 cc<-cor.test(Holz$t01_visperc, Holz$t02_cubes)
@@
 critical.htest(cc)
-@
+```
 
@@
 It can be done by using firstly the 'cor.test' function and then putting the newly created object into the 'critical' function.
-The output will add, underneath the usual output of the 'cor.test' function, the absolute critical value (|rc|).
+The output will add, underneath the usual output of the 'cor.test' function, the absolute critical value (rc).
 The number represents the smallest significant effect.
@@
 
-<<>>=
+```{r}
 tt <- t.test(Holz$t01_visperc [ Holz$female == 1], Holz$t01_visperc [ Holz$female == 2])
@@
 critical.htest(tt)
-@
+```
 
@@
 
-<<>>=
+```{r}
 ll <- lm( t01_visperc ~ ageyr , data = Holz)
 summary(critical(ll))
-@
+```
 
@@
 First, the model needs to be created and saved in an object, then it should be passed into the 'critical' function and wrapped in the 'summary' function.
+
+## Examples with sample size
 
@@
 
-<<>>=
+```{r t from n, echo=FALSE}
 n <- 1200
 critical_t2s(n1 = n , n2 = n , hypothesis = "less")
-@
+```
 
@@
 
-Critical t value from sample size for a paired t-test
+Critical t value from sample size for a paired t-test.
 
-<<>>=
+```{r}
 n <- 15
 critical_t2sp(n = n , hypothesis = "less")
-@
+```
 
@@
 
-<<>>=
+```{r}
 n <- 60
 critical_cor(n = n, hypothesis = "two.sided", test = "z")
-@
+```
 
@@
 
-<<>>=
+```{r}
 n <- 170
 critical_coef(seb = .1, n = n, p = .01, conf.level = .95, hypothesis = "less")
-@
+```
 
 With the function 'critical_coef' one can calculate the critical beta by specifying sample size, standard error of the estimate, p-value, confidence level and hypothesis.
-
-\end{document}

### 2024-06-04 ExPro.Rmd
update

--- 2024-06-04 Examples.Rnw
@@
-\documentclass{article}
+---
+title: "Examples of functions"
+output: html_document
+---
 
-\begin{document}
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
+```
 
-<<>>=
+To install the package the following line of code is needed.
+
+```{r , include = FALSE}
 devtools::load_all()
+```
 
+```{r load, eval=FALSE}
 # require(devtools)
@@
 
-@
+```
 
-<<>>=
+## Examples on real data
+
+Four our examples on real data we used prom the package 'psych' the dataset "holzinger.swineford" which has a series of demografics and scores of different subtests measuring intelligence.
+Once the package is retrieved with the 'library' command, the dataset can be opened using the 'data' function.
+For simplicity we decided to rename it with a shorter name.
+
+```{r open data}
 library(psych)
 library(psychTools)
+```
 
+```{r dataset}
 data("holzinger.swineford")
 Holz <- holzinger.swineford
-@
+```
 
-Critical value for correlation from real data
+#### Critical value for correlation from a cor.test object
 
-<<>>=
-cc<-cor.test(Holz$t01_visperc, Holz$t02_cubes)
+In the following chunk we want to get the critical value for a correlation of two cognitive tests of 301 subjects. We can do so with the 'critical' function:
+
+```{r correlation}
+cc <- cor.test(Holz$t01_visperc, Holz$t02_cubes)
 critical(cc)
-critical.htest(cc)
-@
+```
 
-Critical value for t-test from real data
+First, the 'cor.test' function should be used on the two variables of interest and then the newly created object should be put into the 'critical' function.
+The output will add, underneath the usual output of the 'cor.test' function, the absolute critical value (rc).
+The number represents the smallest significant effect.
 
-<<>>=
+#### Critical value for t-test from real data
+
+Now we want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception, as we assume gender differences.
+In this case, it can be easily done with the same procedure using the 't.test' function:
+
+```{r}
 tt <- t.test(Holz$t01_visperc [ Holz$female == 1], Holz$t01_visperc [ Holz$female == 2])
 critical(tt)
-critical.htest(tt)
-@
+```
 
-Critical value for linear model from real data
+The output now gives a wider range of values: the cohen's d calculated on the data (d), the critical cohen's d (dc), the numerator of the formula for the critical cohen's d (bc), the cohen's d adjusted for small samples (g) and the critical cohen's d adjusted for small samples (gc).
 
-<<>>=
+#### Critical value for linear model from real data
+
+Now we want to calculate the critical coefficients for a linear model with visual perception predicted by age. We can do so by wrapping the 'critical' function into the 'summary' function:
+```{r}
 ll <- lm( t01_visperc ~ ageyr , data = Holz)
 summary(critical(ll))
-@
+```
 
-Critical t value from sample size
+In the output, a column will appear next to the usual column of the Estimates reporting critical values for both the Intercept and the predictor(s).
 
-<<>>=
-n <- 20
+## Examples with sample size
+
+In the next examples we will show the use of the package's functions to calculate critical values in a prospective framework.
+
+#### Critical t value from sample size
+
+Let us say that we want to calculate the critical t value for a t-test for a sample of 1200 subjects for each group, the following chunk shows how to do so with the function 'critical_t2s':
+```{r t from n}
+n <- 1200
 critical_t2s(n1 = n , n2 = n , hypothesis = "less")
-@
+```
 
-Critical t value from sample size for a paired t-test
+The function requires to specify the hypothesis as "less", "greater" or "two.tailed". In output we can find the critical d and the critical g, the adjusted d for small samples.
 
-<<>>=
-critical_t2sp(n = n , hypothesis = "less")
-@
+#### Critical t value from sample size for a paired t-test.
 
-Critical value for a correlation from sample size
+Assuming that we want to calculate the critical value for a paired t-test (i.e. pre-post treatment), we can do so with the function 'critical_t2sp' in the following way:
+```{r}
+n <- 15
+critical_t2sp(n = n , hypothesis = "less", conf.level = .95)
+```
 
-<<>>=
-critical_cor(n = n,hypothesis = "less", test = "z")
-@
+Specifying the direction of the hypothesis and the confidence interval, the function will give as output the standardized and un-standardized critical cohen's d, and the critical cohen's d adjusted for small samples both standardized and un-standardized.
 
-Critical value of a beta from a model
+#### Critical value for a correlation
 
-<<>>=
-critical_coef(seb = .1, n = 20, p = .01, conf.level = .95, hypothesis = "less")
-@
+The same can be done for the critical value for a correlation on a defined sample size with the 'critical_cor' function:
 
-\end{document}
+```{r}
+n <- 60
+critical_cor(n = n, hypothesis = "two.sided", test = "z")
+```
+
+Once more, the direction of the hypothesis and the test to apply, either t-test or z-test, should be specified.
+The output will return the critical correlation value, the degrees of freedom and the type of test used, which should be specified from the user.
+
+#### Critical value of a beta from a model
+
+Let us say that we want to know the critical beta of a model, we can use the 'critical_coef' function as shown in the following chunk:
+
+```{r}
+n <- 170
+critical_coef(seb = .1, n = n, p = .01, conf.level = .95, hypothesis = "less")
+```
+
+It is important to specifying sample size, standard error of the estimate, p-value, confidence level and hypothesis. The output will return the critical beta and the test used.

### 2024-06-05 ExPro.Rmd
updated version

@@
 
-To install the package the following line of code is needed.
+To install the package the following line of code is needed:
 
@@
 
-Four our examples on real data we used prom the package 'psych' the dataset "holzinger.swineford" which has a series of demografics and scores of different subtests measuring intelligence.
-Once the package is retrieved with the 'library' command, the dataset can be opened using the 'data' function.
+For our examples on real data we used from the package 'psych' the dataset "holzinger.swineford" which has a series of demographics and scores of different subtests measuring intelligence on 301 subjects.
+Once the package is retrieved with 'library', the dataset can be opened using 'data("name of the dataset")'.
 For simplicity we decided to rename it with a shorter name.
@@
 
-#### Critical value for correlation from a cor.test object
+#### Critical value for correlation
 
-In the following chunk we want to get the critical value for a correlation of two cognitive tests of 301 subjects. We can do so with the 'critical' function:
+In the following chunk we want to get the critical value for a correlation of two cognitive sub-tests of 301 subjects. We can do so with the 'critical' function:
 
@@
 First, the 'cor.test' function should be used on the two variables of interest and then the newly created object should be put into the 'critical' function.
-The output will add, underneath the usual output of the 'cor.test' function, the absolute critical value (rc).
-The number represents the smallest significant effect.
+The output will add, underneath the usual output of the 'cor.test' function, the absolute critical value (rc), representing the smallest significant effect.
 
-#### Critical value for t-test from real data
+#### Critical value for t-test
 
-Now we want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception, as we assume gender differences.
+Now we want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception.
 In this case, it can be easily done with the same procedure using the 't.test' function:
@@
 
-#### Critical value for linear model from real data
+#### Critical value for linear model
 
@@
 
-In the output, a column will appear next to the usual column of the Estimates reporting critical values for both the Intercept and the predictor(s).
+In the output, a column will appear next to the usual column of the estimates reporting critical values for both the intercept and the predictor(s).
 
@@
 
-The function requires to specify the hypothesis as "less", "greater" or "two.tailed". In output we can find the critical d and the critical g, the adjusted d for small samples.
+The function requires to specify the hypothesis as "less", "greater" or "two.tailed". In output we can find the critical d and the critical g, which is the cohen's d adjusted for small samples.
 
@@
 
-Specifying the direction of the hypothesis and the confidence interval, the function will give as output the standardized and un-standardized critical cohen's d, and the critical cohen's d adjusted for small samples both standardized and un-standardized.
+Specifying the direction of the hypothesis and the confidence interval, the function will give as output the standardized (dzc) and un-standardized (dc) critical cohen's d, and the critical cohen's d adjusted for small samples both standardized (gc) and un-standardized (gzc).
 
-#### Critical value for a correlation
+#### Critical value for a correlation from sample size
 
@@
 
-#### Critical value of a beta from a model
+#### Critical value of a beta from a model from saple size
 

### 2024-08-11 paper/paper.Rmd
adding rmd paper and todo (+ general cleaning)

---
title             : "The benefits of reporting critical effect size values"
shorttitle        : "Reporting critical effect size values"
author: 
  - name          : "Ambra Perugini"
    affiliation   : "1"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Filippo Gambarota"
    affiliation   : "1"
    role:
      - "Writing - Review & Editing"
      - "Supervision"
  - name          : "Enrico Toffalini"
    affiliation   : "2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"
  - name          : "Daniël Lakens"
    affiliation   : "3"
    role:
      - "Writing - Review & Editing"
      - "Supervision"
  - name          : "Massimiliano Pastore"
    affiliation   : "1"
    role:
      - "Writing - Review & Editing"
      - "Supervision"
  - name          : "Livio Finos"
    affiliation   : "4"
    role:
      - "Writing - Review & Editing"
      - "Supervision"
  - name          : "Psicostat"
    affiliation   : "1"
    role:
      - "Writing - Review & Editing"
      - "Supervision"
  - name          : "Gianmarco Altoè"
    affiliation   : "1"
    corresponding: "yes"
    address       : "Via Venezia 8, 35131, Padova, Italy"
    email         : "gianmarco.altoe@unipd.it"
    role:
      - "Writing - Review & Editing"
      - "Supervision"
affiliation:
  - id            : "1"
    institution   : "Department of Developmental and Social Psychology, University of Padova, Italy"
  - id            : "2"
    institution   : "Department of General Psychology, University of Padova, Italy"
  - id: "3"
    institution: "Eindhoven University of Technology, Netherlands"
  - id: "4"
    institution: "Department of Statistics, University of Padova, Italy"

abstract: |
  Scrivere abstract
keywords          : "keywords"
wordcount         : "X"

bibliography      : "bibliography.bib"
csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa7"
floatsintext      : no
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : 
    papaja::apa6_pdf:
      latex_engine: xelatex
header-includes:
  - |
    \makeatletter
    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-1em}%
      {\normalfont\normalsize\bfseries\typesectitle}}

    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-\z@\relax}%
      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
    \makeatother
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "#>")
```

In the present paper, we suggest that researchers systematically report the critical effect size values when they make inference using statistical significance. The "critical effect size values" refers to the smallest statistically significant effect sizes, which depends on the test performed and the characteristics of the sample [@lakens2022sample]. To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with n = 20 participants and uses p < 0.05 as the threshold for statistical significance: in this case, the smallest significant effects (critical values) are Pearson's r = -0.44 and r = 0.44. As we will argue, the critical values can serve as a useful complement to aid in the interpretation of study results, communicate their relevance, help assessing their replicability, particularly when statistical significance is relied upon for inference. This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size or when statistical power may be difficult to determine or deviates from an optimal level.

Despite longstanding criticisms, the Null Hypothesis Significance Testing (NHST) remains a prominent approach for statistical inference in science [@cohen1994; @gigerenzer2004null; @krueger2001null]. Over the past decades, numerous proposals have emerged to enhance inference based on this approach. These include conducting power analyses to determine sample size (n) a priori based on theory-informed target effects, and reporting of effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance [@transue2019apa]. Even for the most skeptical, there are suggestions that NHST may possess merits of its own. For instance, @wilson2020science propose that it could serve as a useful preliminary filter utilized by "original science" for screening potentially interesting effects, that should subsequently be validated by "replication science".

In psychology, power is often severely limited for medium to small effect sizes [@szucs2017empirical], which are generally expected to represent true replicable findings. In fact, resource constraints have been authoritatively listed among acceptable justifications for sample size  [@lakens2022sample], and this scenario is arguably widespread in psychology. However, limited power makes it challenging to distinguish signal from noise, and on average it results in an overestimation of effect sizes when relying on statistical significance for inference [@altoe2020enhancing]. On the other hand, very large samples are occasionally available in psychological science, albeit less frequently. However, as some have argued, excessively large sample sizes risk creating an "everything-is-significant" scenario, where researchers report and discuss practically negligible effects solely because they are statistically significant, regardless of theoretical relevance or whether they just reflect minor procedural artifacts [@wilson2020science]. Thus, while statistical significance can serve as a potentially useful filter, relying solely on it for inference may introduce risks in interpretation. In such cases, we contend that reporting critical values can help put into perspective and convey the relevance of the results.

There are two clear benefits of reporting the critical effect size for a corresponding test. First, when sample sizes are smaller, the critical values inform readers about whether the effect sizes that could lead to a rejection of the null hypothesis are in line with realistic expectations. If the sample size is small, and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes. In these cases, researchers will realized they are not able to collect sufficient data to perform a meaningful test. An a-priori power analysis would typically lead to a similar conclusion, but reporting the critical effect sizes focuses the attention more strongly on which effect sizes are reasonable to expect. Second, in large studies the critical effect sizes will make clear that trivially small effect sizes will be statistically significant. This will focus the attention of researchers on the difference between statistical significance and practical significance. This is especially important in correlational studies on large samples in psychology, where small uncontrolled sources of statistical noise may simultaneously affect and create some shared variance across otherwise unrelated variables, a phenomenon controversially labelled as the 'crud factor' [@orben_crud_2020].

Let us consider the following two scenarios as examples. First, imagine researchers conducting a study involving a between-group comparison. Due to severe resource constraints, they are only able to collect a limited sample size of n = 30 (n = 15 per group). No statistically significant effect is detected with p < 0.05. Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to what extent. Subsequently, they compute the critical values, revealing Cohen's d = -0.75 and d = 0.75. This indicates that any estimated |Cohen's d| < 0.75 will certainly fail to reach statistical significance. By reporting these critical values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance. In essence, this scenario exemplifies a version of the "winner's curse" [@hedges1984estimation], made explicit. In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a "winner") to be associated with an overestimated effect size which will probably be deflated in subsequent replications (i.e., the "curse"), especially if such initial finding was obtained on a small sample. Knowing from the critical values that a specific result must necessarily be associated with an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.

In a second scenario, imagine researchers who gain access to a very large archival dataset (n = 5,000) and decide to explore some bivariate correlations. With such a large sample size, researchers are not constrained by sample size limitations and could instead just focus on effect sizes. However, they opt to legitimately use statistical significance of a two-tailed test as their primary criterion for whether to comment on effects in their further discussions. In this context, the critical effect size values for significance are determined to be correlations of $r = \pm 0.03$. This implies that practically any effect differing from zero, regardless of its small magnitude, may easily attain statistical significance. It is worth noting that such small effects may potentially reflect just minor artifacts [@wilson2020science], such as subtle experimenter effects or slight non-independence among observations, even in meticulously designed studies. While readers may be tempted to interpret any reported effects simply because they attain statistical significance, preliminarily signaling the critical value level can serve as a clear warning that the significance filter might be somehow permissive, thus urging interpretive caution.

A similar scenario may arise in meta-analysis. Despite potential loss of precision due to substantial heterogeneity across effect sizes in different studies, meta-analyses typically synthesize a large amount of evidence, leading even very small average effect sizes to reach significance. While the focus of meta-analysis is generally on estimating effect sizes with uncertainty, statistical significance is routinely reported and interpreted. Thus, in meta-analysis as well, reporting the critical values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.

To provide more tangible illustrations of the practical application of critical values, let us examine two real-world instances drawn from published research. The first one features a small sample, and has been targeted in the series of replications by the Open Science Collaboration [@open2015estimating], while the second one presents with a very large sample and has already attracted attention in the scientific community. The first case pertains to "Study 5" conducted by @mccrea2008self. This experiment compared two groups based on their performance percentage in a math test following a preliminary practice session with feedback of failure. One group was exposed to failure-excusing, self-handicapping thoughts, while both groups underwent the math test afterward. With a modest sample size of 28 participants split into two groups (13 and 15 participants, respectively), a directional one-tailed t-test was conducted on the percentage of correct answers to compare the groups. The resulting effect was statistically significant, $t(26) = 1.87, p (one-tailed) < .05$. Based on the reported data, Cohen's d = 0.736, indicating a considerable effect size. However, the critical value for Cohen's d in this context was 0.645, suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude. Subsequent replication efforts [@open2015estimating] reaffirmed the significance of the effect, and used a somewhat larger sample (n = 61), $t(59)=2.325$. The estimated effect size was somehow lower than that originally reported, however, $d = 0.605$ (here, critical d = 0.428). The second scenario involves the study by @kramer2014experimental, which explored the impact of emotional content on Facebook users' experiences. With a notably large sample size of n = 689'003, the researchers observed a significant effect worsened emotional states when positive posts were reduced, $t(310,044) = −5.63$, $p < 0.001$ (non-directional test). Based on this and other results of a similar, the authors conclude that emotions expressed by other users on Facebook influence our own emotions. Despite the statistical significance, however, the critical values for Cohen's d in this case are -0.006 and +0.006, emphasizing the importance of considering effect size beyond mere significance. This prompts reflection on whether an effect of such magnitude on such a large sample (actual d = 0.02) is truly meaningful in individuals' lives.

# How to Compute Critical Values

In this section, we provide guidance and formulas for computing critical values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's d), correlations (Pearson's r), and raw and standardized coefficients in linear models. These formulas have been incorporated into R functions of the package "criticalESvalues" accessible at: [xxxxx], and are elaborated upon in the subsequent section.

## t-test

For the t-test we considered the two-sample, paired and one-sample tests. As a general approach the $t$ statistics is computed as reported in Equation \@ref(eq:eq-gen-t).

\begin{equation}
    (\#eq:eq-gen-t)
    t = \frac{b}{\text{SE}_{b}}
\end{equation}

Where $b$ is the unstandardized effect size that depends on the type of test. For example, in the two sample t-test or the single mean for a one sample t-test. The denominator is the standard error of the numerator.

Similarly, Equation \@ref(eq:eq-gen-d) formalize a general form of the effect size.

\begin{equation}
    (\#eq:eq-gen-d)
    d = \frac{b}{s}
\end{equation}

Where $b$ is still the unstandardized effect size and $s$ is the standardization term. For example, in the two-sample case, $s$ is the pooled standard deviation between the two samples or the standard deviation of the differences for paired samples.

### two-sample t-test

For the two sample t-test the critical value ($b_c$) is calculated using Equation \@ref(eq:eq-tw-sp). Where $t_c$ is the critical $t$ value calculated using a certain $\alpha$ value (e.g., $0.05$) and $n_1 + n_2 - 2$ degrees of freedom. Then the $b_c$ is simply divided by $s$ obtaining the standardized critical value. For the two-sample case ($s = s_p$ assuming equal variances Equation \@ref(eq:eq-s-p-twsmp) reported the pooled standard deviation.

\begin{equation}
    (\#eq:eq-tw-sp)
    b_c = t_c \times \text{SE}_{b}
\end{equation}

\begin{equation}
    (\#eq:eq-s-p-twsmp)
    s_p = \sqrt{\frac{s_1^2 (n_1 - 1) + s_2^2 (n_2 - 1)}{n_1 + n_2 - 2}}
\end{equation}

When relaxing the assumption of equal variances (i.e., Welch's test) $s$ is simply the square root of the average between the two variances.

### one sample t-test

For the one sample case, the equations are the same. The only difference is that $b$ is the sample mean and $s$ is the sample standard deviation.

### paired sample t-test

For the paired sample case the situation is less straightforward. The reason is that the $t$ statistic is computed using $b$ being the average of the paired differences between the two (paired) samples and $s$ being the standard deviation/error of the differences. While for hypothesis testing and power analysis it is common to calculate the $d$ dividing by the standard deviation of the differences [@lakens2013calculating], this effect size cannot be directly compared to the effect size used in the independent samples t-test, which divides by the pooled standard deviation [@morris2002combining]. Practically, Equation \@ref(eq:eq-sp-sd) depict the relationship between the pooled standard deviation and the standard deviation of the differences.

\begin{equation}
    (\#eq:eq-sp-sd)
    \begin{gathered}
        s_p =  \frac{s_D}{\sqrt{2(1 - \rho)}} \\
        s_D =  s_p \sqrt{2(1 - \rho)}
    \end{gathered}
\end{equation}

Where $\rho$ is the correlation between the two paired samples. Even if the hypothesis testing is computed using $s_D$, we reported the critical value both using the pooled standard deviation and the standard deviation of the differences. Beyond this difference, the equations are still the same as the previous example.

### Hedges's correction

The effect size calculated as in the previous step is known to be inflated thus we calculated also the corrected version usually known as Hedges's $g$ [@Hedges1981-za].

## Correlation Test

For the Pearson's correlation test, the critical value is computed using Equation \@ref(eq:eq-cor-c) where $t_c$ is still the critical value for the $t$ distribution and $n$ is the sample size.

\begin{equation}
    (\#eq:eq-cor-c)
    r_c = \frac{t_c}{\sqrt{n - 2 + t_c^2}}
\end{equation}

Another approach for hypothesis testing of the Pearson's correlation coefficient is using the Fisher's $z$ transformation. Equation \@ref(eq:eq-f-rc) shows how to calculate the critical $F(r_c)$ value where $z_c$ is the critical value of the standard normal distribution with a chosen $\alpha$ level. Then the $F(r_c)$ can be transformed back into a correlation coefficient using Equation \@ref(eq:eq-bk-r).

\begin{equation}
    (\#eq:eq-f-rc)
    \begin{gathered}
        F(r_c) = \frac{z_c}{\sqrt{n - 3}}
    \end{gathered}
\end{equation}

\begin{equation}
    (\#eq:eq-bk-r)
    r_c = \frac{\exp(2z_c)-1}{\exp(2z_c)+1} = \operatorname{tanh}(z_c)
\end{equation}

<!-- %% see better this part -->

## Linear Regression

Hypothesis testing on single coefficients in linear regression (using the \texttt{lm} function in R) is performed using the Wald test. This test is basically a t-test where the $t$ value is calculated dividing the regression parameter $\beta_0, \dots, \beta_j$ by the standard error. Then a $t$ distribution using $n - p - 1$ degrees of freedom where $n$ is the number of observations and $p$ is the number of coefficients beyond the intercept. The $t$ value is calculated as reported in Equation \@ref(eq:eq-wald-lm) and the critical value as reported in Equation \@ref(eq:eq-critical-lm).

\begin{equation}
(\#eq:eq-wald-lm)
    t = \frac{\beta_j}{SE_{\beta_j}}
\end{equation}

\begin{equation}
(\#eq:eq-critical-lm)
    \beta_{jc} = t_c SE_{\beta_j}
\end{equation}

<!-- maybe worth adding more about standardization -->

In this case $\beta_{jc}$ is the unstandardized critical regression coefficient (as $b$ in the t-test section). We implemented also the possibility to use standardized regression coefficients. We can identify two types of standardization namely full and partial. The full standardization involves both the response variable and the predictors. Regression coefficients are interpreted as the increase in standard deviations of the response variable for an increase of one standard deviation in the predictor, keeping fixed all other predictors. The partial standardization involves only the predictors and regression coefficients are interpreted as the increase in the (raw) response variable for a increase in one standard deviation of the predictor. By default, standardizing means dividing a quantity by the standard deviation and eventually centering (i.e., subtracting the mean). Gelman [@Gelman2008-zp] suggested to standardize using two standard deviations because when numeric and binary variables are included the coefficients are not on the same metric. Equation \@ref(eq:eq-critical-lm) can be used but to calculate the critical value for the standardized regression coefficients however the result need to be interpreted according to the chosen approach.

<!-- %% -->

### Meta-analysis

<!-- %% da rivedere -->

Meta-analysis allows to pool information from multiple studies related to specific research question. The main advantage of meta-analysis is pooling multiple studies to obtain a more precise and powerful estimation of the effect. From a statistical point of view, a meta-analysis can be considered as a weighted linear regression with heterogeneous variances. Similarly to standard linear regression, hypothesis testing is performed using Wald $t$ or $z$ tests. The calculation of critical values is the same as reported in Equations \@ref(eq:eq-wald-lm) and \@ref(eq:eq-critical-lm). The only difference is that the standard error of the meta-analysis parameters included both the within-studies variances and eventually the between-studies heterogeneity.

### Other models

Despite we discussed only linear models, the same approach could be theoretically applied to other types of models such as generalized linear models. In fact, we simply need to multiply the critical value of the chosen distribution (e.g., Student $t$ or Standard Normal) by the standard error of the regression coefficient.

# Examples in R

In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package "criticalESvalues" in R. The complete package can be accessed at: [xxxx]. Here, we demonstrate its application through two examples: one example of a t-test on real data and a computation of the critical value for a correlation from sample size. In the Supplementary online materials additional examples can be found for correlation, t-test, paired t-test, linear models and regression coefficients, both from data and from sample size.

First the package should be downloaded and opened with the library function:

```{r, download-package}
# devtools::install_github("filippogambarota/criticalvalue")
library(criticalvalue)
```

For our examples on real data we used from the package 'psych' the dataset "holzinger.swineford" which has a series of demographics and scores of different subtests measuring intelligence on 301 subjects. Once the package is retrieved with 'library', the dataset can be opened using 'data("name of the dataset")'. For simplicity we decided to rename it with a shorter name.

```{r load-dataset}
library(psych)
library(psychTools)
data("holzinger.swineford")
Holz <- holzinger.swineford
```

We want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception. In this case, it can be easily done with the same procedure using the 't.test' function:

```{r t-test}
tt <- t.test(Holz$t01_visperc[Holz$female == 1], 
             Holz$t01_visperc[Holz$female == 2])
critical(tt)
```

The output now gives a wider range of values: the cohen's d calculated on the data (d), the critical cohen's d (dc), the numerator of the formula for the critical cohen's d (bc), the cohen's d adjusted for small samples (g) and the critical cohen's d adjusted for small samples (gc).

In the next example we will show the use of the package's function `critical_cor` to calculate the critical value for a correlation in a prospective framework.

```{r correlation}
n <- 60
critical_cor(n = n, hypothesis = "two.sided", test = "z")
```

The direction of the hypothesis and the test to apply, either t-test or z-test, should be specified. The output will return the critical correlation value, the degrees of freedom and the type of test used.

# Discussion

With the present article, we propose that researchers compute and report the "critical effect size value(s)" in their empirical articles. This is not intended to replace other strategies aimed at enhancing the NHST approach to inference. Such strategies, such as the emphasis on estimating effect sizes with confidence intervals [@transue2019apa] or the a priori planning for statistical power are valuable in their own right. Instead, our proposal serves as a complementary tool, especially beneficial for facilitating the interpretation of results when statistical power deviates from an optimal level (typically falling below, but occasionally exceeding it). Interestingly, the critical value can be retrospectively applied even to already published studies. This potential facilitates potential reframing of the original interpretations. Serving as a tool for retrospective analysis, the critical value may enable a reconsideration of the relevance of previously reported findings.

An advantage of reporting the critical value is that it can be precisely computed in any scenario, without requiring assumptions about the expected effect size, as is the case with power calculations. The critical value represents a directly interpretable benchmark that is especially useful in situations where statistical power is below the desired level and researchers are left otherwise uncertain about how to proceed with the interpretation of a study findings. For example, let us say that we read a published article reporting some effects as statistically significant, while others as not: we suspect that the study may be underpowered, but we are widely uncertain about the magnitude of possible true effects. To what extent can the reported results be interpreted, precisely? Knowing the critical value provides us with a clear benchmark. Conversely, let us say that an effect achieves significance in a very large sample: researchers tend to draw substantive conclusions based on this. But is it of real theoretical relevance? If in comparing two groups, such as controls versus treatment, any Cohen's d > 0.07 would reach significance, is statistical significance enough to signal a "successful" treatment? Maybe yes, even if effects are tiny [e.g., @funder2019evaluating], but knowing the critical value certainly prompts some appropriate interpretive caution.

Reporting the critical value(s) can also be an efficient way to allow researchers to evaluate which findings are statistically significant. For example, in a correlation table researchers customarily add an asterisk to all statistically significant correlations. But as long as all correlations are based on the same sample size, researchers can simply remark 'the critical effect size is *r* = 0.3' and readers will know all correlations larger than this value are statistically significant.

Beyond enhancing study design and statistical inferences based on hypothesis tests, reporting critical value(s) can also serve an educational purpose. It underscores how the distinction between a significant and non-significant result is not solely determined by the presence or absence of a true effect, but also by the sample size. By highlighting a critical value, researchers can become more aware of the possibility of Type 2 errors when results are non-significant. Conversely, in studies with exceedingly large samples and in many meta-analyses, the critical value(s) may serves as a reminder that any observed effect larger than a trivially small value will likely achieve significance. This emphasizes that the mere attainment of statistical significance in a test is not particularly surprising, especially in non-experimental studies.

Real-case scenarios may not always be that simple. Hence, we chose to expand the application of computing critical significance values beyond Cohen's d and correlation to include linear regression with both raw and standardized coefficients. This serves as a first step in computing critical values for a wider array of effects encountered in practical scenarios, where linear models and their extensions are commonly utilized for modeling purposes. A prerequisite is that researchers must be able to identify what parameters in their statistical models reflect the effect sizes of interest, and that they are able to assess their relevance. Notably, however, this prerequisite aligns with the requirements of APA style guidelines concerning the reporting of effect sizes. For further illustration and application, additional examples are provided in the Supplementary online material.

We suggest that reporting critical values is particularly valuable when sample size planning was not feasible or did not occur \textit{a priori}. In cases where optimal power can be attained with a sufficiently large sample size for an effect of a specific magnitude of interest, and this is truly determined \textit{a priori}, the interpretation of both significance and non-significance becomes straightforward. However, when power analysis did not inform the sample size or when power is likely but undeterminedly low, reporting critical values for the obtained sample can help provide context for interpretation. Critical values can be computed and interpreted even retrospectively or for studies that have already been published.

In conclusion, the reporting of the critical value in empirical articles serves as a valuable addition to researchers' toolkit, aimed at augmenting transparency and facilitating the interpretability of their findings. While not designed to supplant existing practices, it provides a useful aid in interpreting newly presented and previously published results, thus advancing the understanding of research outcomes.

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

### 2024-08-11 paper/paper.tex
adding rmd paper and todo (+ general cleaning)

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{keywords\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={The benefits of reporting critical effect size values},
  pdfauthor={Ambra Perugini1, Filippo Gambarota1, Enrico Toffalini2, Daniël Lakens3, Massimiliano Pastore1, Livio Finos4, Psicostat1, \& Gianmarco Altoè1},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{The benefits of reporting critical effect size values}
\author{Ambra Perugini\textsuperscript{1}, Filippo Gambarota\textsuperscript{1}, Enrico Toffalini\textsuperscript{2}, Daniël Lakens\textsuperscript{3}, Massimiliano Pastore\textsuperscript{1}, Livio Finos\textsuperscript{4}, Psicostat\textsuperscript{1}, \& Gianmarco Altoè\textsuperscript{1}}
\date{}

\shorttitle{Reporting critical effect size values}

\authornote{

The authors made the following contributions. Ambra Perugini: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Filippo Gambarota: Writing - Review \& Editing, Supervision; Enrico Toffalini: Writing - Review \& Editing, Supervision; Daniël Lakens: Writing - Review \& Editing, Supervision; Massimiliano Pastore: Writing - Review \& Editing, Supervision; Livio Finos: Writing - Review \& Editing, Supervision; Psicostat: Writing - Review \& Editing, Supervision; Gianmarco Altoè: Writing - Review \& Editing, Supervision.

Correspondence concerning this article should be addressed to Gianmarco Altoè, Via Venezia 8, 35131, Padova, Italy. E-mail: \href{mailto:gianmarco.altoe@unipd.it}{\nolinkurl{gianmarco.altoe@unipd.it}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Developmental and Social Psychology, University of Padova, Italy\\\textsuperscript{2} Department of General Psychology, University of Padova, Italy\\\textsuperscript{3} Eindhoven University of Technology, Netherlands\\\textsuperscript{4} Department of Statistics, University of Padova, Italy}

\abstract{%
Scrivere abstract
}

\begin{document}
\maketitle

In the present paper, we suggest that researchers systematically report the critical effect size values when they make inference using statistical significance. The ``critical effect size values'' refers to the smallest statistically significant effect sizes, which depends on the test performed and the characteristics of the sample (Lakens, 2022). To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with n = 20 participants and uses p \textless{} 0.05 as the threshold for statistical significance: in this case, the smallest significant effects (critical values) are Pearson's r = -0.44 and r = 0.44. As we will argue, the critical values can serve as a useful complement to aid in the interpretation of study results, communicate their relevance, help assessing their replicability, particularly when statistical significance is relied upon for inference. This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size or when statistical power may be difficult to determine or deviates from an optimal level.

Despite longstanding criticisms, the Null Hypothesis Significance Testing (NHST) remains a prominent approach for statistical inference in science (Cohen, 1994; Gigerenzer et al., 2004; Krueger, 2001). Over the past decades, numerous proposals have emerged to enhance inference based on this approach. These include conducting power analyses to determine sample size (n) a priori based on theory-informed target effects, and reporting of effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance (Transue, 2019). Even for the most skeptical, there are suggestions that NHST may possess merits of its own. For instance, Wilson et al. (2020) propose that it could serve as a useful preliminary filter utilized by ``original science'' for screening potentially interesting effects, that should subsequently be validated by ``replication science''.

In psychology, power is often severely limited for medium to small effect sizes (Szucs \& Ioannidis, 2017), which are generally expected to represent true replicable findings. In fact, resource constraints have been authoritatively listed among acceptable justifications for sample size (Lakens, 2022), and this scenario is arguably widespread in psychology. However, limited power makes it challenging to distinguish signal from noise, and on average it results in an overestimation of effect sizes when relying on statistical significance for inference (Altoè et al., 2020). On the other hand, very large samples are occasionally available in psychological science, albeit less frequently. However, as some have argued, excessively large sample sizes risk creating an ``everything-is-significant'' scenario, where researchers report and discuss practically negligible effects solely because they are statistically significant, regardless of theoretical relevance or whether they just reflect minor procedural artifacts (Wilson et al., 2020). Thus, while statistical significance can serve as a potentially useful filter, relying solely on it for inference may introduce risks in interpretation. In such cases, we contend that reporting critical values can help put into perspective and convey the relevance of the results.

There are two clear benefits of reporting the critical effect size for a corresponding test. First, when sample sizes are smaller, the critical values inform readers about whether the effect sizes that could lead to a rejection of the null hypothesis are in line with realistic expectations. If the sample size is small, and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes. In these cases, researchers will realized they are not able to collect sufficient data to perform a meaningful test. An a-priori power analysis would typically lead to a similar conclusion, but reporting the critical effect sizes focuses the attention more strongly on which effect sizes are reasonable to expect. Second, in large studies the critical effect sizes will make clear that trivially small effect sizes will be statistically significant. This will focus the attention of researchers on the difference between statistical significance and practical significance. This is especially important in correlational studies on large samples in psychology, where small uncontrolled sources of statistical noise may simultaneously affect and create some shared variance across otherwise unrelated variables, a phenomenon controversially labelled as the `crud factor' (Orben \& Lakens, 2020).

Let us consider the following two scenarios as examples. First, imagine researchers conducting a study involving a between-group comparison. Due to severe resource constraints, they are only able to collect a limited sample size of n = 30 (n = 15 per group). No statistically significant effect is detected with p \textless{} 0.05. Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to what extent. Subsequently, they compute the critical values, revealing Cohen's d = -0.75 and d = 0.75. This indicates that any estimated \textbar Cohen's d\textbar{} \textless{} 0.75 will certainly fail to reach statistical significance. By reporting these critical values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance. In essence, this scenario exemplifies a version of the ``winner's curse'' (Hedges, 1984), made explicit. In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a ``winner'') to be associated with an overestimated effect size which will probably be deflated in subsequent replications (i.e., the ``curse''), especially if such initial finding was obtained on a small sample. Knowing from the critical values that a specific result must necessarily be associated with an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.

In a second scenario, imagine researchers who gain access to a very large archival dataset (n = 5,000) and decide to explore some bivariate correlations. With such a large sample size, researchers are not constrained by sample size limitations and could instead just focus on effect sizes. However, they opt to legitimately use statistical significance of a two-tailed test as their primary criterion for whether to comment on effects in their further discussions. In this context, the critical effect size values for significance are determined to be correlations of \(r = \pm 0.03\). This implies that practically any effect differing from zero, regardless of its small magnitude, may easily attain statistical significance. It is worth noting that such small effects may potentially reflect just minor artifacts (Wilson et al., 2020), such as subtle experimenter effects or slight non-independence among observations, even in meticulously designed studies. While readers may be tempted to interpret any reported effects simply because they attain statistical significance, preliminarily signaling the critical value level can serve as a clear warning that the significance filter might be somehow permissive, thus urging interpretive caution.

A similar scenario may arise in meta-analysis. Despite potential loss of precision due to substantial heterogeneity across effect sizes in different studies, meta-analyses typically synthesize a large amount of evidence, leading even very small average effect sizes to reach significance. While the focus of meta-analysis is generally on estimating effect sizes with uncertainty, statistical significance is routinely reported and interpreted. Thus, in meta-analysis as well, reporting the critical values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.

To provide more tangible illustrations of the practical application of critical values, let us examine two real-world instances drawn from published research. The first one features a small sample, and has been targeted in the series of replications by the Open Science Collaboration (Collaboration, 2015), while the second one presents with a very large sample and has already attracted attention in the scientific community. The first case pertains to ``Study 5'' conducted by McCrea (2008). This experiment compared two groups based on their performance percentage in a math test following a preliminary practice session with feedback of failure. One group was exposed to failure-excusing, self-handicapping thoughts, while both groups underwent the math test afterward. With a modest sample size of 28 participants split into two groups (13 and 15 participants, respectively), a directional one-tailed t-test was conducted on the percentage of correct answers to compare the groups. The resulting effect was statistically significant, \(t(26) = 1.87, p (one-tailed) < .05\). Based on the reported data, Cohen's d = 0.736, indicating a considerable effect size. However, the critical value for Cohen's d in this context was 0.645, suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude. Subsequent replication efforts (Collaboration, 2015) reaffirmed the significance of the effect, and used a somewhat larger sample (n = 61), \(t(59)=2.325\). The estimated effect size was somehow lower than that originally reported, however, \(d = 0.605\) (here, critical d = 0.428). The second scenario involves the study by Kramer et al. (2014), which explored the impact of emotional content on Facebook users' experiences. With a notably large sample size of n = 689'003, the researchers observed a significant effect worsened emotional states when positive posts were reduced, \(t(310,044) = −5.63\), \(p < 0.001\) (non-directional test). Based on this and other results of a similar, the authors conclude that emotions expressed by other users on Facebook influence our own emotions. Despite the statistical significance, however, the critical values for Cohen's d in this case are -0.006 and +0.006, emphasizing the importance of considering effect size beyond mere significance. This prompts reflection on whether an effect of such magnitude on such a large sample (actual d = 0.02) is truly meaningful in individuals' lives.

\section{How to Compute Critical Values}\label{how-to-compute-critical-values}

In this section, we provide guidance and formulas for computing critical values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's d), correlations (Pearson's r), and raw and standardized coefficients in linear models. These formulas have been incorporated into R functions of the package ``criticalESvalues'' accessible at: {[}xxxxx{]}, and are elaborated upon in the subsequent section.

\subsection{t-test}\label{t-test}

For the t-test we considered the two-sample, paired and one-sample tests. As a general approach the \(t\) statistics is computed as reported in Equation \eqref{eq:eq-gen-t}.

\begin{equation}
    \label{eq:eq-gen-t}
    t = \frac{b}{\text{SE}_{b}}
\end{equation}

Where \(b\) is the unstandardized effect size that depends on the type of test. For example, in the two sample t-test or the single mean for a one sample t-test. The denominator is the standard error of the numerator.

Similarly, Equation \eqref{eq:eq-gen-d} formalize a general form of the effect size.

\begin{equation}
    \label{eq:eq-gen-d}
    d = \frac{b}{s}
\end{equation}

Where \(b\) is still the unstandardized effect size and \(s\) is the standardization term. For example, in the two-sample case, \(s\) is the pooled standard deviation between the two samples or the standard deviation of the differences for paired samples.

\subsubsection{two-sample t-test}\label{two-sample-t-test}

For the two sample t-test the critical value (\(b_c\)) is calculated using Equation \eqref{eq:eq-tw-sp}. Where \(t_c\) is the critical \(t\) value calculated using a certain \(\alpha\) value (e.g., \(0.05\)) and \(n_1 + n_2 - 2\) degrees of freedom. Then the \(b_c\) is simply divided by \(s\) obtaining the standardized critical value. For the two-sample case (\(s = s_p\) assuming equal variances Equation \eqref{eq:eq-s-p-twsmp} reported the pooled standard deviation.

\begin{equation}
    \label{eq:eq-tw-sp}
    b_c = t_c \times \text{SE}_{b}
\end{equation}

\begin{equation}
    \label{eq:eq-s-p-twsmp}
    s_p = \sqrt{\frac{s_1^2 (n_1 - 1) + s_2^2 (n_2 - 1)}{n_1 + n_2 - 2}}
\end{equation}

When relaxing the assumption of equal variances (i.e., Welch's test) \(s\) is simply the square root of the average between the two variances.

\subsubsection{one sample t-test}\label{one-sample-t-test}

For the one sample case, the equations are the same. The only difference is that \(b\) is the sample mean and \(s\) is the sample standard deviation.

\subsubsection{paired sample t-test}\label{paired-sample-t-test}

For the paired sample case the situation is less straightforward. The reason is that the \(t\) statistic is computed using \(b\) being the average of the paired differences between the two (paired) samples and \(s\) being the standard deviation/error of the differences. While for hypothesis testing and power analysis it is common to calculate the \(d\) dividing by the standard deviation of the differences (Lakens, 2013), this effect size cannot be directly compared to the effect size used in the independent samples t-test, which divides by the pooled standard deviation (Morris \& DeShon, 2002). Practically, Equation \eqref{eq:eq-sp-sd} depict the relationship between the pooled standard deviation and the standard deviation of the differences.

\begin{equation}
    \label{eq:eq-sp-sd}
    \begin{gathered}
        s_p =  \frac{s_D}{\sqrt{2(1 - \rho)}} \\
        s_D =  s_p \sqrt{2(1 - \rho)}
    \end{gathered}
\end{equation}

Where \(\rho\) is the correlation between the two paired samples. Even if the hypothesis testing is computed using \(s_D\), we reported the critical value both using the pooled standard deviation and the standard deviation of the differences. Beyond this difference, the equations are still the same as the previous example.

\subsubsection{Hedges's correction}\label{hedgess-correction}

The effect size calculated as in the previous step is known to be inflated thus we calculated also the corrected version usually known as Hedges's \(g\) (Hedges, 1981).

\subsection{Correlation Test}\label{correlation-test}

For the Pearson's correlation test, the critical value is computed using Equation \eqref{eq:eq-cor-c} where \(t_c\) is still the critical value for the \(t\) distribution and \(n\) is the sample size.

\begin{equation}
    \label{eq:eq-cor-c}
    r_c = \frac{t_c}{\sqrt{n - 2 + t_c^2}}
\end{equation}

Another approach for hypothesis testing of the Pearson's correlation coefficient is using the Fisher's \(z\) transformation. Equation \eqref{eq:eq-f-rc} shows how to calculate the critical \(F(r_c)\) value where \(z_c\) is the critical value of the standard normal distribution with a chosen \(\alpha\) level. Then the \(F(r_c)\) can be transformed back into a correlation coefficient using Equation \eqref{eq:eq-bk-r}.

\begin{equation}
    \label{eq:eq-f-rc}
    \begin{gathered}
        F(r_c) = \frac{z_c}{\sqrt{n - 3}}
    \end{gathered}
\end{equation}

\begin{equation}
    \label{eq:eq-bk-r}
    r_c = \frac{\exp(2z_c)-1}{\exp(2z_c)+1} = \operatorname{tanh}(z_c)
\end{equation}

\subsection{Linear Regression}\label{linear-regression}

Hypothesis testing on single coefficients in linear regression (using the \texttt{lm} function in R) is performed using the Wald test. This test is basically a t-test where the \(t\) value is calculated dividing the regression parameter \(\beta_0, \dots, \beta_j\) by the standard error. Then a \(t\) distribution using \(n - p - 1\) degrees of freedom where \(n\) is the number of observations and \(p\) is the number of coefficients beyond the intercept. The \(t\) value is calculated as reported in Equation \eqref{eq:eq-wald-lm} and the critical value as reported in Equation \eqref{eq:eq-critical-lm}.

\begin{equation}
\label{eq:eq-wald-lm}
    t = \frac{\beta_j}{SE_{\beta_j}}
\end{equation}

\begin{equation}
\label{eq:eq-critical-lm}
    \beta_{jc} = t_c SE_{\beta_j}
\end{equation}

In this case \(\beta_{jc}\) is the unstandardized critical regression coefficient (as \(b\) in the t-test section). We implemented also the possibility to use standardized regression coefficients. We can identify two types of standardization namely full and partial. The full standardization involves both the response variable and the predictors. Regression coefficients are interpreted as the increase in standard deviations of the response variable for an increase of one standard deviation in the predictor, keeping fixed all other predictors. The partial standardization involves only the predictors and regression coefficients are interpreted as the increase in the (raw) response variable for a increase in one standard deviation of the predictor. By default, standardizing means dividing a quantity by the standard deviation and eventually centering (i.e., subtracting the mean). Gelman (Gelman, 2008) suggested to standardize using two standard deviations because when numeric and binary variables are included the coefficients are not on the same metric. Equation \eqref{eq:eq-critical-lm} can be used but to calculate the critical value for the standardized regression coefficients however the result need to be interpreted according to the chosen approach.

\subsubsection{Meta-analysis}\label{meta-analysis}

Meta-analysis allows to pool information from multiple studies related to specific research question. The main advantage of meta-analysis is pooling multiple studies to obtain a more precise and powerful estimation of the effect. From a statistical point of view, a meta-analysis can be considered as a weighted linear regression with heterogeneous variances. Similarly to standard linear regression, hypothesis testing is performed using Wald \(t\) or \(z\) tests. The calculation of critical values is the same as reported in Equations \eqref{eq:eq-wald-lm} and \eqref{eq:eq-critical-lm}. The only difference is that the standard error of the meta-analysis parameters included both the within-studies variances and eventually the between-studies heterogeneity.

\subsubsection{Other models}\label{other-models}

Despite we discussed only linear models, the same approach could be theoretically applied to other types of models such as generalized linear models. In fact, we simply need to multiply the critical value of the chosen distribution (e.g., Student \(t\) or Standard Normal) by the standard error of the regression coefficient.

\section{Examples in R}\label{examples-in-r}

In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package ``criticalESvalues'' in R. The complete package can be accessed at: {[}xxxx{]}. Here, we demonstrate its application through two examples: one example of a t-test on real data and a computation of the critical value for a correlation from sample size. In the Supplementary online materials additional examples can be found for correlation, t-test, paired t-test, linear models and regression coefficients, both from data and from sample size.

First the package should be downloaded and opened with the library function:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# devtools::install\_github("filippogambarota/criticalvalue")}
\FunctionTok{library}\NormalTok{(criticalvalue)}
\end{Highlighting}
\end{Shaded}

For our examples on real data we used from the package `psych' the dataset ``holzinger.swineford'' which has a series of demographics and scores of different subtests measuring intelligence on 301 subjects. Once the package is retrieved with `library', the dataset can be opened using `data(``name of the dataset'')'. For simplicity we decided to rename it with a shorter name.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(psych)}
\FunctionTok{library}\NormalTok{(psychTools)}
\FunctionTok{data}\NormalTok{(}\StringTok{"holzinger.swineford"}\NormalTok{)}
\NormalTok{Holz }\OtherTok{\textless{}{-}}\NormalTok{ holzinger.swineford}
\end{Highlighting}
\end{Shaded}

We want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception. In this case, it can be easily done with the same procedure using the `t.test' function:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tt }\OtherTok{\textless{}{-}} \FunctionTok{t.test}\NormalTok{(Holz}\SpecialCharTok{$}\NormalTok{t01\_visperc[Holz}\SpecialCharTok{$}\NormalTok{female }\SpecialCharTok{==} \DecValTok{1}\NormalTok{], }
\NormalTok{             Holz}\SpecialCharTok{$}\NormalTok{t01\_visperc[Holz}\SpecialCharTok{$}\NormalTok{female }\SpecialCharTok{==} \DecValTok{2}\NormalTok{])}
\FunctionTok{critical}\NormalTok{(tt)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> 
#>  Welch Two Sample t-test
#> 
#> data:  Holz$t01_visperc[Holz$female == 1] and Holz$t01_visperc[Holz$female == 2]
#> t = 1.4095, df = 298.9, p-value = 0.1597
#> alternative hypothesis: true difference in means is not equal to 0
#> 95 percent confidence interval:
#>  -0.06419229  0.38823118
#> sample estimates:
#> mean of x mean of y 
#>  4.314090  4.152071 
#> 
#> |== Effect Size and Critical Value ==| 
#> d = 0.1623665 dc = ± 0.2269609 bc = ± 0.2262117 
#> g = 0.1619587 gc = ± 0.2263909
\end{verbatim}

The output now gives a wider range of values: the cohen's d calculated on the data (d), the critical cohen's d (dc), the numerator of the formula for the critical cohen's d (bc), the cohen's d adjusted for small samples (g) and the critical cohen's d adjusted for small samples (gc).

In the next example we will show the use of the package's function \texttt{critical\_cor} to calculate the critical value for a correlation in a prospective framework.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{n }\OtherTok{\textless{}{-}} \DecValTok{60}
\FunctionTok{critical\_cor}\NormalTok{(}\AttributeTok{n =}\NormalTok{ n, }\AttributeTok{hypothesis =} \StringTok{"two.sided"}\NormalTok{, }\AttributeTok{test =} \StringTok{"z"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> $rc
#> [1] 0.2539247
#> 
#> $df
#> [1] 58
#> 
#> $test
#> [1] "z"
\end{verbatim}

The direction of the hypothesis and the test to apply, either t-test or z-test, should be specified. The output will return the critical correlation value, the degrees of freedom and the type of test used.

\section{Discussion}\label{discussion}

With the present article, we propose that researchers compute and report the ``critical effect size value(s)'' in their empirical articles. This is not intended to replace other strategies aimed at enhancing the NHST approach to inference. Such strategies, such as the emphasis on estimating effect sizes with confidence intervals (Transue, 2019) or the a priori planning for statistical power are valuable in their own right. Instead, our proposal serves as a complementary tool, especially beneficial for facilitating the interpretation of results when statistical power deviates from an optimal level (typically falling below, but occasionally exceeding it). Interestingly, the critical value can be retrospectively applied even to already published studies. This potential facilitates potential reframing of the original interpretations. Serving as a tool for retrospective analysis, the critical value may enable a reconsideration of the relevance of previously reported findings.

An advantage of reporting the critical value is that it can be precisely computed in any scenario, without requiring assumptions about the expected effect size, as is the case with power calculations. The critical value represents a directly interpretable benchmark that is especially useful in situations where statistical power is below the desired level and researchers are left otherwise uncertain about how to proceed with the interpretation of a study findings. For example, let us say that we read a published article reporting some effects as statistically significant, while others as not: we suspect that the study may be underpowered, but we are widely uncertain about the magnitude of possible true effects. To what extent can the reported results be interpreted, precisely? Knowing the critical value provides us with a clear benchmark. Conversely, let us say that an effect achieves significance in a very large sample: researchers tend to draw substantive conclusions based on this. But is it of real theoretical relevance? If in comparing two groups, such as controls versus treatment, any Cohen's d \textgreater{} 0.07 would reach significance, is statistical significance enough to signal a ``successful'' treatment? Maybe yes, even if effects are tiny (e.g., Funder \& Ozer, 2019), but knowing the critical value certainly prompts some appropriate interpretive caution.

Reporting the critical value(s) can also be an efficient way to allow researchers to evaluate which findings are statistically significant. For example, in a correlation table researchers customarily add an asterisk to all statistically significant correlations. But as long as all correlations are based on the same sample size, researchers can simply remark `the critical effect size is \emph{r} = 0.3' and readers will know all correlations larger than this value are statistically significant.

Beyond enhancing study design and statistical inferences based on hypothesis tests, reporting critical value(s) can also serve an educational purpose. It underscores how the distinction between a significant and non-significant result is not solely determined by the presence or absence of a true effect, but also by the sample size. By highlighting a critical value, researchers can become more aware of the possibility of Type 2 errors when results are non-significant. Conversely, in studies with exceedingly large samples and in many meta-analyses, the critical value(s) may serves as a reminder that any observed effect larger than a trivially small value will likely achieve significance. This emphasizes that the mere attainment of statistical significance in a test is not particularly surprising, especially in non-experimental studies.

Real-case scenarios may not always be that simple. Hence, we chose to expand the application of computing critical significance values beyond Cohen's d and correlation to include linear regression with both raw and standardized coefficients. This serves as a first step in computing critical values for a wider array of effects encountered in practical scenarios, where linear models and their extensions are commonly utilized for modeling purposes. A prerequisite is that researchers must be able to identify what parameters in their statistical models reflect the effect sizes of interest, and that they are able to assess their relevance. Notably, however, this prerequisite aligns with the requirements of APA style guidelines concerning the reporting of effect sizes. For further illustration and application, additional examples are provided in the Supplementary online material.

We suggest that reporting critical values is particularly valuable when sample size planning was not feasible or did not occur \textit{a priori}. In cases where optimal power can be attained with a sufficiently large sample size for an effect of a specific magnitude of interest, and this is truly determined \textit{a priori}, the interpretation of both significance and non-significance becomes straightforward. However, when power analysis did not inform the sample size or when power is likely but undeterminedly low, reporting critical values for the obtained sample can help provide context for interpretation. Critical values can be computed and interpreted even retrospectively or for studies that have already been published.

In conclusion, the reporting of the critical value in empirical articles serves as a valuable addition to researchers' toolkit, aimed at augmenting transparency and facilitating the interpretability of their findings. While not designed to supplant existing practices, it provides a useful aid in interpreting newly presented and previously published results, thus advancing the understanding of research outcomes.

\newpage

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-altoe2020enhancing}
Altoè, G., Bertoldo, G., Zandonella Callegher, C., Toffalini, E., Calcagnı̀, A., Finos, L., \& Pastore, M. (2020). Enhancing statistical inference in psychological research via prospective and retrospective design analysis. \emph{Frontiers in Psychology}, \emph{10}, 499756.

\bibitem[\citeproctext]{ref-cohen1994}
Cohen, J. (1994). The earth is round (p \textless{} .05). \emph{American Psychologist}, \emph{49}(12), 997--1003.

\bibitem[\citeproctext]{ref-open2015estimating}
Collaboration, O. S. (2015). Estimating the reproducibility of psychological science. \emph{Science}, \emph{349}(6251), aac4716.

\bibitem[\citeproctext]{ref-funder2019evaluating}
Funder, D. C., \& Ozer, D. J. (2019). Evaluating effect size in psychological research: Sense and nonsense. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168.

\bibitem[\citeproctext]{ref-Gelman2008-zp}
Gelman, A. (2008). Scaling regression inputs by dividing by two standard deviations. \emph{Statistics in Medicine}, \emph{27}, 2865--2873. \url{https://doi.org/10.1002/sim.3107}

\bibitem[\citeproctext]{ref-gigerenzer2004null}
Gigerenzer, G., Krauss, S., \& Vitouch, O. (2004). The null ritual. \emph{The Sage Handbook of Quantitative Methodology for the Social Sciences}, 391--408.

\bibitem[\citeproctext]{ref-Hedges1981-za}
Hedges, L. V. (1981). Distribution theory for glass's estimator of effect size and related estimators. \emph{Journal of Educational and Behavioral Statistics: A Quarterly Publication Sponsored by the American Educational Research Association and the American Statistical Association}, \emph{6}, 107--128. \url{https://doi.org/10.3102/10769986006002107}

\bibitem[\citeproctext]{ref-hedges1984estimation}
Hedges, L. V. (1984). Estimation of effect size under nonrandom sampling: The effects of censoring studies yielding statistically insignificant mean differences. \emph{Journal of Educational Statistics}, \emph{9}(1), 61--85.

\bibitem[\citeproctext]{ref-kramer2014experimental}
Kramer, A. D., Guillory, J. E., \& Hancock, J. T. (2014). Experimental evidence of massive-scale emotional contagion through social networks. \emph{Proceedings of the National Academy of Sciences of the United States of America}, \emph{111}(24), 8788.

\bibitem[\citeproctext]{ref-krueger2001null}
Krueger, J. (2001). Null hypothesis significance testing: On the survival of a flawed method. \emph{American Psychologist}, \emph{56}(1), 16.

\bibitem[\citeproctext]{ref-lakens2013calculating}
Lakens, D. (2013). Calculating and reporting effect sizes to facilitate cumulative science: A practical primer for t-tests and ANOVAs. \emph{Frontiers in Psychology}, \emph{4}, 62627.

\bibitem[\citeproctext]{ref-lakens2022sample}
Lakens, D. (2022). Sample size justification. \emph{Collabra: Psychology}, \emph{8}(1), 33267.

\bibitem[\citeproctext]{ref-mccrea2008self}
McCrea, S. M. (2008). Self-handicapping, excuse making, and counterfactual thinking: Consequences for self-esteem and future motivation. \emph{Journal of Personality and Social Psychology}, \emph{95}(2), 274.

\bibitem[\citeproctext]{ref-morris2002combining}
Morris, S. B., \& DeShon, R. P. (2002). Combining effect size estimates in meta-analysis with repeated measures and independent-groups designs. \emph{Psychological Methods}, \emph{7}(1), 105.

\bibitem[\citeproctext]{ref-orben_crud_2020}
Orben, A., \& Lakens, D. (2020). Crud ({Re}){Defined}. \emph{Advances in Methods and Practices in Psychological Science}, \emph{3}(2), 238--247. \url{https://doi.org/10.1177/2515245920917961}

\bibitem[\citeproctext]{ref-szucs2017empirical}
Szucs, D., \& Ioannidis, J. P. (2017). Empirical assessment of published effect sizes and power in the recent cognitive neuroscience and psychology literature. \emph{PLoS Biology}, \emph{15}(3), e2000797.

\bibitem[\citeproctext]{ref-transue2019apa}
Transue, B. (2019). \emph{APA style 7th edition}.

\bibitem[\citeproctext]{ref-wilson2020science}
Wilson, B. M., Harris, C. R., \& Wixted, J. T. (2020). Science is not a signal detection problem. \emph{Proceedings of the National Academy of Sciences}, \emph{117}(11), 5559--5567.

\end{CSLReferences}

\end{document}

### 2024-08-12 paper/paper.Rmd
^^

@@
 The "critical effect size values" refers to the smallest statistically significant effect sizes, which depends on the test performed and the characteristics of the sample [@lakens2022sample].
-To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with n = 20 participants and uses p < 0.05 as the threshold for statistical significance: in this case, the smallest significant effects (critical values) are Pearson's r = -0.44 and r = 0.44.
+To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with $n = 20$ participants and uses $p < 0.05$ as the threshold for statistical significance: in this case, the smallest significant effects (critical values) are Pearson's $r = -0.44$ and $r = 0.44$.
 As we will argue, the critical values can serve as a useful complement to aid in the interpretation of study results, communicate their relevance, help assessing their replicability, particularly when statistical significance is relied upon for inference.
@@
 Over the past decades, numerous proposals have emerged to enhance inference based on this approach.
-These include conducting power analyses to determine sample size (n) a priori based on theory-informed target effects, and reporting of effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance [@transue2019apa].
+These include conducting power analyses to determine sample size ($n$) a priori based on theory-informed target effects, and reporting of effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance [@transue2019apa].
 Even for the most skeptical, there are suggestions that NHST may possess merits of its own.
@@
 First, imagine researchers conducting a study involving a between-group comparison.
-Due to severe resource constraints, they are only able to collect a limited sample size of n = 30 (n = 15 per group).
-No statistically significant effect is detected with p < 0.05.
+Due to severe resource constraints, they are only able to collect a limited sample size of $n = 30$ ($n = 15$ per group).
+No statistically significant effect is detected with $p < 0.05$.
 Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to what extent.
-Subsequently, they compute the critical values, revealing Cohen's d = -0.75 and d = 0.75.
-This indicates that any estimated |Cohen's d| < 0.75 will certainly fail to reach statistical significance.
+Subsequently, they compute the critical values, revealing Cohen's $d = -0.75$ and $d = 0.75$.
+This indicates that any estimated |Cohen's $d| < 0.75$ will certainly fail to reach statistical significance.
 By reporting these critical values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance.
@@
 
-In a second scenario, imagine researchers who gain access to a very large archival dataset (n = 5,000) and decide to explore some bivariate correlations.
+In a second scenario, imagine researchers who gain access to a very large archival dataset ($n = 5,000$) and decide to explore some bivariate correlations.
 With such a large sample size, researchers are not constrained by sample size limitations and could instead just focus on effect sizes.
@@
 The resulting effect was statistically significant, $t(26) = 1.87, p (one-tailed) < .05$.
-Based on the reported data, Cohen's d = 0.736, indicating a considerable effect size.
-However, the critical value for Cohen's d in this context was 0.645, suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude.
-Subsequent replication efforts [@open2015estimating] reaffirmed the significance of the effect, and used a somewhat larger sample (n = 61), $t(59)=2.325$.
+Based on the reported data, Cohen's $d = 0.736$, indicating a considerable effect size.
+However, the critical value for Cohen's $d$ in this context was $0.645$, suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude.
+Subsequent replication efforts [@open2015estimating] reaffirmed the significance of the effect, and used a somewhat larger sample ($n = 61$), $t(59)=2.325$.
 The estimated effect size was somehow lower than that originally reported, however, $d = 0.605$ (here, critical d = 0.428).
 The second scenario involves the study by @kramer2014experimental, which explored the impact of emotional content on Facebook users' experiences.
-With a notably large sample size of n = 689'003, the researchers observed a significant effect worsened emotional states when positive posts were reduced, $t(310,044) = −5.63$, $p < 0.001$ (non-directional test).
+With a notably large sample size of $n = 689'003$, the researchers observed a significant effect worsened emotional states when positive posts were reduced, $t(310,044) = −5.63$, $p < 0.001$ (non-directional test).
 Based on this and other results of a similar, the authors conclude that emotions expressed by other users on Facebook influence our own emotions.
-Despite the statistical significance, however, the critical values for Cohen's d in this case are -0.006 and +0.006, emphasizing the importance of considering effect size beyond mere significance.
-This prompts reflection on whether an effect of such magnitude on such a large sample (actual d = 0.02) is truly meaningful in individuals' lives.
+Despite the statistical significance, however, the critical values for Cohen's $d$ in this case are $-0.006$ and $+0.006$, emphasizing the importance of considering effect size beyond mere significance.
+This prompts reflection on whether an effect of such magnitude on such a large sample (actual $d = 0.02$) is truly meaningful in individuals' lives.
 
@@
 
-In this section, we provide guidance and formulas for computing critical values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's d), correlations (Pearson's r), and raw and standardized coefficients in linear models.
+In this section, we provide guidance and formulas for computing critical values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's $d$), correlations (Pearson's r), and raw and standardized coefficients in linear models.
 These formulas have been incorporated into R functions of the package "criticalESvalues" accessible at: [xxxxx], and are elaborated upon in the subsequent section.
@@
 
-The output now gives a wider range of values: the cohen's d calculated on the data (d), the critical cohen's d (dc), the numerator of the formula for the critical cohen's d (bc), the cohen's d adjusted for small samples (g) and the critical cohen's d adjusted for small samples (gc).
+The output now gives a wider range of values: the Cohen's $d$ calculated on the data ($d$), the critical Cohen's $d$ ($dc$), the numerator of the formula for the critical Cohen's $d$ ($bc$), the Cohen's $d$ adjusted for small samples ($g$) and the critical Cohen's $d$ adjusted for small samples ($gc$).
 
@@
 
-The direction of the hypothesis and the test to apply, either t-test or z-test, should be specified.
+The direction of the hypothesis and the test to apply, either $t$-test or $z$-test, should be specified.
 The output will return the critical correlation value, the degrees of freedom and the type of test used.
@@
 But is it of real theoretical relevance?
-If in comparing two groups, such as controls versus treatment, any Cohen's d > 0.07 would reach significance, is statistical significance enough to signal a "successful" treatment?
+If in comparing two groups, such as controls versus treatment, any Cohen's $d > 0.07$ would reach significance, is statistical significance enough to signal a "successful" treatment?
 Maybe yes, even if effects are tiny [e.g., @funder2019evaluating], but knowing the critical value certainly prompts some appropriate interpretive caution.
@@
 For example, in a correlation table researchers customarily add an asterisk to all statistically significant correlations.
-But as long as all correlations are based on the same sample size, researchers can simply remark 'the critical effect size is *r* = 0.3' and readers will know all correlations larger than this value are statistically significant.
+But as long as all correlations are based on the same sample size, researchers can simply remark 'the critical effect size is $r = 0.3$' and readers will know all correlations larger than this value are statistically significant.
 
@@
 Real-case scenarios may not always be that simple.
-Hence, we chose to expand the application of computing critical significance values beyond Cohen's d and correlation to include linear regression with both raw and standardized coefficients.
+Hence, we chose to expand the application of computing critical significance values beyond Cohen's $d$ and correlation to include linear regression with both raw and standardized coefficients.
 This serves as a first step in computing critical values for a wider array of effects encountered in practical scenarios, where linear models and their extensions are commonly utilized for modeling purposes.

### 2024-08-12 paper/paper.tex
^^

@@
 \fi
-% definitions for citeproc citations
-\NewDocumentCommand\citeproctext{}{}
-\NewDocumentCommand\citeproc{mm}{%
-  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
-\makeatletter
- % allow citations to break across lines
- \let\@cite@ofmt\@firstofone
- % avoid brackets around text for \cite:
- \def\@biblabel#1{}
- \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
-\makeatother
 \newlength{\cslhangindent}
@@
 \setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
- {\begin{list}{}{%
-  \setlength{\itemindent}{0pt}
-  \setlength{\leftmargin}{0pt}
-  \setlength{\parsep}{0pt}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
   % turn on hanging indent if param 1 is 1
   \ifodd #1
-   \setlength{\leftmargin}{\cslhangindent}
-   \setlength{\itemindent}{-1\cslhangindent}
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
   \fi
   % set entry spacing
-  \setlength{\itemsep}{#2\baselineskip}}}
- {\end{list}}
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
 \usepackage{calc}
-\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 
-% \usepackage{etoolbox}
 \makeatletter
-\patchcmd{\HyOrg@maketitle}
+\usepackage{etoolbox}
+\patchcmd{\maketitle}
   {\section{\normalfont\normalsize\abstractname}}
@@
   {}{\typeout{Failed to patch abstract.}}
-\patchcmd{\HyOrg@maketitle}
+\patchcmd{\maketitle}
   {\section{\protect\normalfont{\@title}}}
@@
 \fi
-\usepackage{bookmark}
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 The ``critical effect size values'' refers to the smallest statistically significant effect sizes, which depends on the test performed and the characteristics of the sample (Lakens, 2022).
-To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with n = 20 participants and uses p \textless{} 0.05 as the threshold for statistical significance: in this case, the smallest significant effects (critical values) are Pearson's r = -0.44 and r = 0.44.
+To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with \(n = 20\) participants and uses \(p < 0.05\) as the threshold for statistical significance: in this case, the smallest significant effects (critical values) are Pearson's \(r = -0.44\) and \(r = 0.44\).
 As we will argue, the critical values can serve as a useful complement to aid in the interpretation of study results, communicate their relevance, help assessing their replicability, particularly when statistical significance is relied upon for inference.
@@
 Over the past decades, numerous proposals have emerged to enhance inference based on this approach.
-These include conducting power analyses to determine sample size (n) a priori based on theory-informed target effects, and reporting of effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance (Transue, 2019).
+These include conducting power analyses to determine sample size (\(n\)) a priori based on theory-informed target effects, and reporting of effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance (Transue, 2019).
 Even for the most skeptical, there are suggestions that NHST may possess merits of its own.
@@
 First, imagine researchers conducting a study involving a between-group comparison.
-Due to severe resource constraints, they are only able to collect a limited sample size of n = 30 (n = 15 per group).
-No statistically significant effect is detected with p \textless{} 0.05.
+Due to severe resource constraints, they are only able to collect a limited sample size of \(n = 30\) (\(n = 15\) per group).
+No statistically significant effect is detected with \(p < 0.05\).
 Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to what extent.
-Subsequently, they compute the critical values, revealing Cohen's d = -0.75 and d = 0.75.
-This indicates that any estimated \textbar Cohen's d\textbar{} \textless{} 0.75 will certainly fail to reach statistical significance.
+Subsequently, they compute the critical values, revealing Cohen's \(d = -0.75\) and \(d = 0.75\).
+This indicates that any estimated \textbar Cohen's \(d| < 0.75\) will certainly fail to reach statistical significance.
 By reporting these critical values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance.
@@
 
-In a second scenario, imagine researchers who gain access to a very large archival dataset (n = 5,000) and decide to explore some bivariate correlations.
+In a second scenario, imagine researchers who gain access to a very large archival dataset (\(n = 5,000\)) and decide to explore some bivariate correlations.
 With such a large sample size, researchers are not constrained by sample size limitations and could instead just focus on effect sizes.
@@
 The resulting effect was statistically significant, \(t(26) = 1.87, p (one-tailed) < .05\).
-Based on the reported data, Cohen's d = 0.736, indicating a considerable effect size.
-However, the critical value for Cohen's d in this context was 0.645, suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude.
-Subsequent replication efforts (Collaboration, 2015) reaffirmed the significance of the effect, and used a somewhat larger sample (n = 61), \(t(59)=2.325\).
+Based on the reported data, Cohen's \(d = 0.736\), indicating a considerable effect size.
+However, the critical value for Cohen's d in this context was \(0.645\), suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude.
+Subsequent replication efforts (Collaboration, 2015) reaffirmed the significance of the effect, and used a somewhat larger sample (\(n = 61\)), \(t(59)=2.325\).
 The estimated effect size was somehow lower than that originally reported, however, \(d = 0.605\) (here, critical d = 0.428).
@@
 (2014), which explored the impact of emotional content on Facebook users' experiences.
-With a notably large sample size of n = 689'003, the researchers observed a significant effect worsened emotional states when positive posts were reduced, \(t(310,044) = −5.63\), \(p < 0.001\) (non-directional test).
+With a notably large sample size of \(n = 689'003\), the researchers observed a significant effect worsened emotional states when positive posts were reduced, \(t(310,044) = −5.63\), \(p < 0.001\) (non-directional test).
 Based on this and other results of a similar, the authors conclude that emotions expressed by other users on Facebook influence our own emotions.
-Despite the statistical significance, however, the critical values for Cohen's d in this case are -0.006 and +0.006, emphasizing the importance of considering effect size beyond mere significance.
-This prompts reflection on whether an effect of such magnitude on such a large sample (actual d = 0.02) is truly meaningful in individuals' lives.
+Despite the statistical significance, however, the critical values for Cohen's d in this case are \(-0.006\) and \(+0.006\), emphasizing the importance of considering effect size beyond mere significance.
+This prompts reflection on whether an effect of such magnitude on such a large sample (actual \(d = 0.02\)) is truly meaningful in individuals' lives.
 
-\section{How to Compute Critical Values}\label{how-to-compute-critical-values}
+\hypertarget{how-to-compute-critical-values}{%
+\section{How to Compute Critical Values}\label{how-to-compute-critical-values}}
 
@@
 
-\subsection{t-test}\label{t-test}
+\hypertarget{t-test}{%
+\subsection{t-test}\label{t-test}}
 
@@
 
-\subsubsection{two-sample t-test}\label{two-sample-t-test}
+\hypertarget{two-sample-t-test}{%
+\subsubsection{two-sample t-test}\label{two-sample-t-test}}
 
@@
 
-\subsubsection{one sample t-test}\label{one-sample-t-test}
+\hypertarget{one-sample-t-test}{%
+\subsubsection{one sample t-test}\label{one-sample-t-test}}
 
@@
 
-\subsubsection{paired sample t-test}\label{paired-sample-t-test}
+\hypertarget{paired-sample-t-test}{%
+\subsubsection{paired sample t-test}\label{paired-sample-t-test}}
 
@@
 
-\subsubsection{Hedges's correction}\label{hedgess-correction}
+\hypertarget{hedgess-correction}{%
+\subsubsection{Hedges's correction}\label{hedgess-correction}}
 
@@
 
-\subsection{Correlation Test}\label{correlation-test}
+\hypertarget{correlation-test}{%
+\subsection{Correlation Test}\label{correlation-test}}
 
@@
 
-\subsection{Linear Regression}\label{linear-regression}
+\hypertarget{linear-regression}{%
+\subsection{Linear Regression}\label{linear-regression}}
 
@@
 
-\subsubsection{Meta-analysis}\label{meta-analysis}
+\hypertarget{meta-analysis}{%
+\subsubsection{Meta-analysis}\label{meta-analysis}}
 
@@
 
-\subsubsection{Other models}\label{other-models}
+\hypertarget{other-models}{%
+\subsubsection{Other models}\label{other-models}}
 
@@
 
-\section{Examples in R}\label{examples-in-r}
+\hypertarget{examples-in-r}{%
+\section{Examples in R}\label{examples-in-r}}
 
@@
 
-The output now gives a wider range of values: the cohen's d calculated on the data (d), the critical cohen's d (dc), the numerator of the formula for the critical cohen's d (bc), the cohen's d adjusted for small samples (g) and the critical cohen's d adjusted for small samples (gc).
+The output now gives a wider range of values: the Cohen's d calculated on the data (\(d\)), the critical Cohen's d (\(dc\)), the numerator of the formula for the critical Cohen's d (\(bc\)), the Cohen's d adjusted for small samples (\(g\)) and the critical Cohen's d adjusted for small samples (\(gc\)).
 
@@
 
-The direction of the hypothesis and the test to apply, either t-test or z-test, should be specified.
+The direction of the hypothesis and the test to apply, either \(t\)-test or \(z\)-test, should be specified.
 The output will return the critical correlation value, the degrees of freedom and the type of test used.
 
-\section{Discussion}\label{discussion}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
@@
 But is it of real theoretical relevance?
-If in comparing two groups, such as controls versus treatment, any Cohen's d \textgreater{} 0.07 would reach significance, is statistical significance enough to signal a ``successful'' treatment?
+If in comparing two groups, such as controls versus treatment, any Cohen's \(d > 0.07\) would reach significance, is statistical significance enough to signal a ``successful'' treatment?
 Maybe yes, even if effects are tiny (e.g., Funder \& Ozer, 2019), but knowing the critical value certainly prompts some appropriate interpretive caution.
@@
 For example, in a correlation table researchers customarily add an asterisk to all statistically significant correlations.
-But as long as all correlations are based on the same sample size, researchers can simply remark `the critical effect size is \emph{r} = 0.3' and readers will know all correlations larger than this value are statistically significant.
+But as long as all correlations are based on the same sample size, researchers can simply remark `the critical effect size is \(r = 0.3\)' and readers will know all correlations larger than this value are statistically significant.
 
@@
 
-\section{References}\label{references}
+\hypertarget{references}{%
+\section{References}\label{references}}
 
-\phantomsection\label{refs}
+\hypertarget{refs}{}
 \begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-altoe2020enhancing}
+\leavevmode\vadjust pre{\hypertarget{ref-altoe2020enhancing}{}}%
 Altoè, G., Bertoldo, G., Zandonella Callegher, C., Toffalini, E., Calcagnı̀, A., Finos, L., \& Pastore, M.
@@
 
-\bibitem[\citeproctext]{ref-cohen1994}
+\leavevmode\vadjust pre{\hypertarget{ref-cohen1994}{}}%
 Cohen, J. (1994). The earth is round (p \textless{} .05). \emph{American Psychologist}, \emph{49}(12), 997--1003.
 
-\bibitem[\citeproctext]{ref-open2015estimating}
+\leavevmode\vadjust pre{\hypertarget{ref-open2015estimating}{}}%
 Collaboration, O. S. (2015). Estimating the reproducibility of psychological science. \emph{Science}, \emph{349}(6251), aac4716.
 
-\bibitem[\citeproctext]{ref-funder2019evaluating}
+\leavevmode\vadjust pre{\hypertarget{ref-funder2019evaluating}{}}%
 Funder, D. C., \& Ozer, D. J. (2019). Evaluating effect size in psychological research: Sense and nonsense. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168.
 
-\bibitem[\citeproctext]{ref-Gelman2008-zp}
+\leavevmode\vadjust pre{\hypertarget{ref-Gelman2008-zp}{}}%
 Gelman, A. (2008). Scaling regression inputs by dividing by two standard deviations. \emph{Statistics in Medicine}, \emph{27}, 2865--2873. \url{https://doi.org/10.1002/sim.3107}
 
-\bibitem[\citeproctext]{ref-gigerenzer2004null}
+\leavevmode\vadjust pre{\hypertarget{ref-gigerenzer2004null}{}}%
 Gigerenzer, G., Krauss, S., \& Vitouch, O. (2004). The null ritual. \emph{The Sage Handbook of Quantitative Methodology for the Social Sciences}, 391--408.
 
-\bibitem[\citeproctext]{ref-Hedges1981-za}
+\leavevmode\vadjust pre{\hypertarget{ref-Hedges1981-za}{}}%
 Hedges, L.
@@
 
-\bibitem[\citeproctext]{ref-hedges1984estimation}
+\leavevmode\vadjust pre{\hypertarget{ref-hedges1984estimation}{}}%
 Hedges, L.
@@
 
-\bibitem[\citeproctext]{ref-kramer2014experimental}
+\leavevmode\vadjust pre{\hypertarget{ref-kramer2014experimental}{}}%
 Kramer, A.
@@
 
-\bibitem[\citeproctext]{ref-krueger2001null}
+\leavevmode\vadjust pre{\hypertarget{ref-krueger2001null}{}}%
 Krueger, J. (2001). Null hypothesis significance testing: On the survival of a flawed method. \emph{American Psychologist}, \emph{56}(1), 16.
 
-\bibitem[\citeproctext]{ref-lakens2013calculating}
+\leavevmode\vadjust pre{\hypertarget{ref-lakens2013calculating}{}}%
 Lakens, D. (2013). Calculating and reporting effect sizes to facilitate cumulative science: A practical primer for t-tests and ANOVAs. \emph{Frontiers in Psychology}, \emph{4}, 62627.
 
-\bibitem[\citeproctext]{ref-lakens2022sample}
+\leavevmode\vadjust pre{\hypertarget{ref-lakens2022sample}{}}%
 Lakens, D. (2022). Sample size justification. \emph{Collabra: Psychology}, \emph{8}(1), 33267.
 
-\bibitem[\citeproctext]{ref-mccrea2008self}
+\leavevmode\vadjust pre{\hypertarget{ref-mccrea2008self}{}}%
 McCrea, S.
@@
 
-\bibitem[\citeproctext]{ref-morris2002combining}
+\leavevmode\vadjust pre{\hypertarget{ref-morris2002combining}{}}%
 Morris, S. B., \& DeShon, R. P. (2002). Combining effect size estimates in meta-analysis with repeated measures and independent-groups designs. \emph{Psychological Methods}, \emph{7}(1), 105.
 
-\bibitem[\citeproctext]{ref-orben_crud_2020}
+\leavevmode\vadjust pre{\hypertarget{ref-orben_crud_2020}{}}%
 Orben, A., \& Lakens, D. (2020). Crud ({Re}){Defined}. \emph{Advances in Methods and Practices in Psychological Science}, \emph{3}(2), 238--247. \url{https://doi.org/10.1177/2515245920917961}
 
-\bibitem[\citeproctext]{ref-szucs2017empirical}
+\leavevmode\vadjust pre{\hypertarget{ref-szucs2017empirical}{}}%
 Szucs, D., \& Ioannidis, J.
@@
 
-\bibitem[\citeproctext]{ref-transue2019apa}
+\leavevmode\vadjust pre{\hypertarget{ref-transue2019apa}{}}%
 Transue, B. (2019). \emph{APA style 7th edition}.
 
-\bibitem[\citeproctext]{ref-wilson2020science}
+\leavevmode\vadjust pre{\hypertarget{ref-wilson2020science}{}}%
 Wilson, B. M., Harris, C. R., \& Wixted, J. T. (2020). Science is not a signal detection problem. \emph{Proceedings of the National Academy of Sciences}, \emph{117}(11), 5559--5567.

### 2024-08-14 paper/paper.Rmd
added abstract and keywords

@@
 abstract: |
-  Scrivere abstract
-keywords          : "keywords"
+  Critical effect size values represent the smallest detectable effect that would reach statistical significance given a certain sample size.
+These values are highly applicable in various scenarios, as they do not require specifying a plausible effect size of interest.
+Reporting critical effect size values may be especially useful when adopting NHST (Null Hypothesis Significance Testing) and the sample size has not been planned a priori, or when there is uncertainty on the expected effect size.
+To assist researchers in calculating critical values, we have developed an R package that allows to report the critical effect size values of model parameters.
+The package's functions are implemented for group comparisons, correlations, linear regression, and meta-analysis.
+What we propose could benefit researchers during the planning phase of the study by helping them to understand the limitations and strengths of their research design, and also as a retrospective tool to potentially reframe original interpretations.
+
+keywords          : critical effect size values, statistical significance, fucntions implementation 
 wordcount         : "X"

### 2024-08-14 paper/paper.tex
added abstract and keywords

@@
 {}{\PatchFailed}
-\keywords{keywords\newline\indent Word count: X}
+\keywords{critical effect size values, statistical significance, fucntions implementation\newline\indent Word count: X}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
   pdflang={en-EN},
-  pdfkeywords={keywords},
+  pdfkeywords={critical effect size values, statistical significance, fucntions implementation},
   hidelinks,
@@
 \abstract{%
-Scrivere abstract
+Critical effect size values represent the smallest detectable effect that would reach statistical significance given a certain sample size.
+These values are highly applicable in various scenarios, as they do not require specifying a plausible effect size of interest.
+Reporting critical effect size values may be especially useful when adopting NHST (Null Hypothesis Significance Testing) and the sample size has not been planned a priori, or when there is uncertainty on the expected effect size.
+To assist researchers in calculating critical values, we have developed an R package that allows to report the critical effect size values of model parameters.
+The package's functions are implemented for group comparisons, correlations, linear regression, and meta-analysis.
+What we propose could benefit researchers during the planning phase of the study by helping them to understand the limitations and strengths of their research design, and also as a retrospective tool to potentially reframe original interpretations.
 }
@@
 Based on the reported data, Cohen's \(d = 0.736\), indicating a considerable effect size.
-However, the critical value for Cohen's d in this context was \(0.645\), suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude.
+However, the critical value for Cohen's \(d\) in this context was \(0.645\), suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude.
 Subsequent replication efforts (Collaboration, 2015) reaffirmed the significance of the effect, and used a somewhat larger sample (\(n = 61\)), \(t(59)=2.325\).
@@
 Based on this and other results of a similar, the authors conclude that emotions expressed by other users on Facebook influence our own emotions.
-Despite the statistical significance, however, the critical values for Cohen's d in this case are \(-0.006\) and \(+0.006\), emphasizing the importance of considering effect size beyond mere significance.
+Despite the statistical significance, however, the critical values for Cohen's \(d\) in this case are \(-0.006\) and \(+0.006\), emphasizing the importance of considering effect size beyond mere significance.
 This prompts reflection on whether an effect of such magnitude on such a large sample (actual \(d = 0.02\)) is truly meaningful in individuals' lives.
@@
 
-In this section, we provide guidance and formulas for computing critical values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's d), correlations (Pearson's r), and raw and standardized coefficients in linear models.
+In this section, we provide guidance and formulas for computing critical values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's r), and raw and standardized coefficients in linear models.
 These formulas have been incorporated into R functions of the package ``criticalESvalues'' accessible at: {[}xxxxx{]}, and are elaborated upon in the subsequent section.
@@
 
-The output now gives a wider range of values: the Cohen's d calculated on the data (\(d\)), the critical Cohen's d (\(dc\)), the numerator of the formula for the critical Cohen's d (\(bc\)), the Cohen's d adjusted for small samples (\(g\)) and the critical Cohen's d adjusted for small samples (\(gc\)).
+The output now gives a wider range of values: the Cohen's \(d\) calculated on the data (\(d\)), the critical Cohen's \(d\) (\(dc\)), the numerator of the formula for the critical Cohen's \(d\) (\(bc\)), the Cohen's \(d\) adjusted for small samples (\(g\)) and the critical Cohen's \(d\) adjusted for small samples (\(gc\)).
 
@@
 Real-case scenarios may not always be that simple.
-Hence, we chose to expand the application of computing critical significance values beyond Cohen's d and correlation to include linear regression with both raw and standardized coefficients.
+Hence, we chose to expand the application of computing critical significance values beyond Cohen's \(d\) and correlation to include linear regression with both raw and standardized coefficients.
 This serves as a first step in computing critical values for a wider array of effects encountered in practical scenarios, where linear models and their extensions are commonly utilized for modeling purposes.

### 2024-08-19 paper/paper.Rmd
m

@@
 First, when sample sizes are smaller, the critical values inform readers about whether the effect sizes that could lead to a rejection of the null hypothesis are in line with realistic expectations.
-If the sample size is small, and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes.
-In these cases, researchers will realized they are not able to collect sufficient data to perform a meaningful test.
+If the sample size is small, and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, in these cases, researchers will realized they are not able to collect sufficient data to perform a meaningful test.
 An a-priori power analysis would typically lead to a similar conclusion, but reporting the critical effect sizes focuses the attention more strongly on which effect sizes are reasonable to expect.
@@
 No statistically significant effect is detected with $p < 0.05$.
-Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to what extent.
+Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to some extent.
 Subsequently, they compute the critical values, revealing Cohen's $d = -0.75$ and $d = 0.75$.
@@
 With a notably large sample size of $n = 689'003$, the researchers observed a significant effect worsened emotional states when positive posts were reduced, $t(310,044) = −5.63$, $p < 0.001$ (non-directional test).
-Based on this and other results of a similar, the authors conclude that emotions expressed by other users on Facebook influence our own emotions.
+Based on this and other results, the authors conclude that emotions expressed by other users on Facebook influence our own emotions.
 Despite the statistical significance, however, the critical values for Cohen's $d$ in this case are $-0.006$ and $+0.006$, emphasizing the importance of considering effect size beyond mere significance.
@@
 
-# How to Compute Critical Values
+# How to Compute Critical Effect Size Values
 
@@
 Then the $b_c$ is simply divided by $s$ obtaining the standardized critical value.
-For the two-sample case ($s = s_p$ assuming equal variances Equation \@ref(eq:eq-s-p-twsmp) reported the pooled standard deviation.
+For the two-sample case ($s = s_p$ assuming equal variances) Equation \@ref(eq:eq-s-p-twsmp) reported the pooled standard deviation.
 
@@
 While for hypothesis testing and power analysis it is common to calculate the $d$ dividing by the standard deviation of the differences [@lakens2013calculating], this effect size cannot be directly compared to the effect size used in the independent samples t-test, which divides by the pooled standard deviation [@morris2002combining].
-Practically, Equation \@ref(eq:eq-sp-sd) depict the relationship between the pooled standard deviation and the standard deviation of the differences.
+Practically, Equation \@ref(eq:eq-sp-sd) depicts the relationship between the pooled standard deviation and the standard deviation of the differences.
 
@@
 
-In this case $\beta_{jc}$ is the unstandardized critical regression coefficient (as $b$ in the t-test section).
+In this case $\beta_{jc}$ is the unstandardized critical regression coefficient (as $b$ in the $t$-test section).
 We implemented also the possibility to use standardized regression coefficients.
-We can identify two types of standardization namely full and partial.
+We can identify two types of standardization, namely full and partial.
 The full standardization involves both the response variable and the predictors.
@@
 Gelman [@Gelman2008-zp] suggested to standardize using two standard deviations because when numeric and binary variables are included the coefficients are not on the same metric.
-Equation \@ref(eq:eq-critical-lm) can be used but to calculate the critical value for the standardized regression coefficients however the result need to be interpreted according to the chosen approach.
+Equation \@ref(eq:eq-critical-lm) can be used but to calculate the critical value for the standardized regression coefficients the result need to be interpreted according to the chosen approach.
 
@@
 
-Meta-analysis allows to pool information from multiple studies related to specific research question.
+Meta-analysis allows to pool information from multiple studies related to a specific research question.
 The main advantage of meta-analysis is pooling multiple studies to obtain a more precise and powerful estimation of the effect.
@@
 
-The output now gives a wider range of values: the Cohen's $d$ calculated on the data ($d$), the critical Cohen's $d$ ($dc$), the numerator of the formula for the critical Cohen's $d$ ($bc$), the Cohen's $d$ adjusted for small samples ($g$) and the critical Cohen's $d$ adjusted for small samples ($gc$).
+The output gives a wide range of values: the Cohen's $d$ calculated on the data ($d$), the critical Cohen's $d$ ($dc$), the numerator of the formula for the critical Cohen's $d$ ($bc$), the Cohen's $d$ adjusted for small samples ($g$) and the critical Cohen's $d$ adjusted for small samples ($gc$).
 
@@
 Interestingly, the critical value can be retrospectively applied even to already published studies.
-This potential facilitates potential reframing of the original interpretations.
+This possible application facilitates potential reframing of the original interpretations.
 Serving as a tool for retrospective analysis, the critical value may enable a reconsideration of the relevance of previously reported findings.
@@
 For example, in a correlation table researchers customarily add an asterisk to all statistically significant correlations.
-But as long as all correlations are based on the same sample size, researchers can simply remark 'the critical effect size is $r = 0.3$' and readers will know all correlations larger than this value are statistically significant.
+But as long as all correlations are based on the same sample size, researchers can simply remark 'the critical effect size is $r = 0.3$' and readers will know that all correlations larger than this value are statistically significant.
 
@@
 
-We suggest that reporting critical values is particularly valuable when sample size planning was not feasible or did not occur \textit{a priori}.
+We suggest that reporting critical effect size values is particularly valuable when sample size planning was not feasible or did not occur \textit{a priori}.
 In cases where optimal power can be attained with a sufficiently large sample size for an effect of a specific magnitude of interest, and this is truly determined \textit{a priori}, the interpretation of both significance and non-significance becomes straightforward.

### 2024-08-19 paper/paper.Rmd
m

@@
     role:
+      - "Writing – Original Draft Preparation"
       - "Writing - Review & Editing"
-      - "Supervision"
+      - "Methodology"
   - name          : "Enrico Toffalini"
@@
     role:
+      - "Writing – Original Draft Preparation"
       - "Writing - Review & Editing"
@@
     role:
+      - "Conceptualization"
       - "Writing - Review & Editing"

### 2024-08-19 paper/paper.tex
m

@@
 The authors made the following contributions.
-Ambra Perugini: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Filippo Gambarota: Writing - Review \& Editing, Supervision; Enrico Toffalini: Writing - Review \& Editing, Supervision; Daniël Lakens: Writing - Review \& Editing, Supervision; Massimiliano Pastore: Writing - Review \& Editing, Supervision; Livio Finos: Writing - Review \& Editing, Supervision; Psicostat: Writing - Review \& Editing, Supervision; Gianmarco Altoè: Writing - Review \& Editing, Supervision.
+Ambra Perugini: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Filippo Gambarota: Writing -- Original Draft Preparation, Writing - Review \& Editing, Methodology; Enrico Toffalini: Writing -- Original Draft Preparation, Writing - Review \& Editing, Supervision; Daniël Lakens: Conceptualization, Writing - Review \& Editing, Supervision; Massimiliano Pastore: Writing - Review \& Editing, Supervision; Livio Finos: Writing - Review \& Editing, Supervision; Psicostat: Writing - Review \& Editing, Supervision; Gianmarco Altoè: Conceptualization, Writing - Review \& Editing, Supervision.
 
@@
 No statistically significant effect is detected with \(p < 0.05\).
-Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to what extent.
+Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to some extent.
 Subsequently, they compute the critical values, revealing Cohen's \(d = -0.75\) and \(d = 0.75\).
@@
 With a notably large sample size of \(n = 689'003\), the researchers observed a significant effect worsened emotional states when positive posts were reduced, \(t(310,044) = −5.63\), \(p < 0.001\) (non-directional test).
-Based on this and other results of a similar, the authors conclude that emotions expressed by other users on Facebook influence our own emotions.
+Based on this and other results, the authors conclude that emotions expressed by other users on Facebook influence our own emotions.
 Despite the statistical significance, however, the critical values for Cohen's \(d\) in this case are \(-0.006\) and \(+0.006\), emphasizing the importance of considering effect size beyond mere significance.
@@
 
-\hypertarget{how-to-compute-critical-values}{%
-\section{How to Compute Critical Values}\label{how-to-compute-critical-values}}
+\hypertarget{how-to-compute-critical-effect-size-values}{%
+\section{How to Compute Critical Effect Size Values}\label{how-to-compute-critical-effect-size-values}}
 
@@
 Then the \(b_c\) is simply divided by \(s\) obtaining the standardized critical value.
-For the two-sample case (\(s = s_p\) assuming equal variances Equation \eqref{eq:eq-s-p-twsmp} reported the pooled standard deviation.
+For the two-sample case (\(s = s_p\) assuming equal variances) Equation \eqref{eq:eq-s-p-twsmp} reported the pooled standard deviation.
 
@@
 While for hypothesis testing and power analysis it is common to calculate the \(d\) dividing by the standard deviation of the differences (Lakens, 2013), this effect size cannot be directly compared to the effect size used in the independent samples t-test, which divides by the pooled standard deviation (Morris \& DeShon, 2002).
-Practically, Equation \eqref{eq:eq-sp-sd} depict the relationship between the pooled standard deviation and the standard deviation of the differences.
+Practically, Equation \eqref{eq:eq-sp-sd} depicts the relationship between the pooled standard deviation and the standard deviation of the differences.
 
@@
 
-In this case \(\beta_{jc}\) is the unstandardized critical regression coefficient (as \(b\) in the t-test section).
+In this case \(\beta_{jc}\) is the unstandardized critical regression coefficient (as \(b\) in the \(t\)-test section).
 We implemented also the possibility to use standardized regression coefficients.
-We can identify two types of standardization namely full and partial.
+We can identify two types of standardization, namely full and partial.
 The full standardization involves both the response variable and the predictors.
@@
 Gelman (Gelman, 2008) suggested to standardize using two standard deviations because when numeric and binary variables are included the coefficients are not on the same metric.
-Equation \eqref{eq:eq-critical-lm} can be used but to calculate the critical value for the standardized regression coefficients however the result need to be interpreted according to the chosen approach.
+Equation \eqref{eq:eq-critical-lm} can be used but to calculate the critical value for the standardized regression coefficients the result need to be interpreted according to the chosen approach.
 
@@
 
-The output now gives a wider range of values: the Cohen's \(d\) calculated on the data (\(d\)), the critical Cohen's \(d\) (\(dc\)), the numerator of the formula for the critical Cohen's \(d\) (\(bc\)), the Cohen's \(d\) adjusted for small samples (\(g\)) and the critical Cohen's \(d\) adjusted for small samples (\(gc\)).
+The output gives a wide range of values: the Cohen's \(d\) calculated on the data (\(d\)), the critical Cohen's \(d\) (\(dc\)), the numerator of the formula for the critical Cohen's \(d\) (\(bc\)), the Cohen's \(d\) adjusted for small samples (\(g\)) and the critical Cohen's \(d\) adjusted for small samples (\(gc\)).
 
@@
 For example, in a correlation table researchers customarily add an asterisk to all statistically significant correlations.
-But as long as all correlations are based on the same sample size, researchers can simply remark `the critical effect size is \(r = 0.3\)' and readers will know all correlations larger than this value are statistically significant.
+But as long as all correlations are based on the same sample size, researchers can simply remark `the critical effect size is \(r = 0.3\)' and readers will know that all correlations larger than this value are statistically significant.
 

### 2024-08-20 paper/paper.Rmd
m

--- 2024-08-19 paper/paper.Rmd
@@
       - "Writing - Review & Editing"
+      - "Software"
   - name          : "Filippo Gambarota"
@@
     role:
+      - "Writing – Original Draft Preparation"
       - "Writing - Review & Editing"
-      - "Supervision"
+      - "Methodology"
+      - "Software"
   - name          : "Enrico Toffalini"
@@
     role:
+      - "Writing – Original Draft Preparation"
       - "Writing - Review & Editing"
+      - "Software"
       - "Supervision"
@@
 There are two clear benefits of reporting the critical effect size for a corresponding test.
-First, when sample sizes are smaller, the critical values inform readers about whether the effect sizes that could lead to a rejection of the null hypothesis are in line with realistic expectations.
+First, when sample sizes are small, the critical values inform readers about whether the effect sizes that could lead to a rejection of the null hypothesis are in line with realistic expectations.
 If the sample size is small, and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, in these cases, researchers will realized they are not able to collect sufficient data to perform a meaningful test.
@@
 In this section, we provide guidance and formulas for computing critical values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's $d$), correlations (Pearson's r), and raw and standardized coefficients in linear models.
-These formulas have been incorporated into R functions of the package "criticalESvalues" accessible at: [xxxxx], and are elaborated upon in the subsequent section.
+These formulas have been incorporated into R functions of the package "criticalvalue" accessible at: https://github.com/filippogambarota/criticalvalue, and are elaborated upon in the subsequent section.
 
@@
 
-In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package "criticalESvalues" in R.
-The complete package can be accessed at: [xxxx].
-Here, we demonstrate its application through two examples: one example of a t-test on real data and a computation of the critical value for a correlation from sample size.
-In the Supplementary online materials additional examples can be found for correlation, t-test, paired t-test, linear models and regression coefficients, both from data and from sample size.
+In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package "criticalvalue" in R.
+The complete package can be accessed at: https://github.com/filippogambarota/criticalvalue.
+Here, we demonstrate its application through two examples: one example of a t-test on real data and a computation of the critical effect size values for a correlation from sample size.
+It is important to note that, in general, depending on the researcher's hypotheses, the package allows for the calculation of either two critical effect size values for two-tailed tests or a single critical effect size value for one-tailed tests, for which the direction must be specified.
+In the Supplementary online materials additional examples can be found for correlation, t-test, paired t-test, linear models and regression coefficients.
 
@@
 
-We want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception.
-In this case, it can be easily done with the same procedure using the 't.test' function:
+We want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception. In this case, it can be easily done using the 't.test' function:
 
@@
 tt <- t.test(Holz$t01_visperc[Holz$female == 1], 
-             Holz$t01_visperc[Holz$female == 2])
+             Holz$t01_visperc[Holz$female == 2], var.equal = T)
 critical(tt)
@@
 The output gives a wide range of values: the Cohen's $d$ calculated on the data ($d$), the critical Cohen's $d$ ($dc$), the numerator of the formula for the critical Cohen's $d$ ($bc$), the Cohen's $d$ adjusted for small samples ($g$) and the critical Cohen's $d$ adjusted for small samples ($gc$).
+The variance is set to be equal, but if that is not the case of your data, you can also run Welch two sample $t$-test and obtain the critical effect size values for that.
 
-In the next example we will show the use of the package's function `critical_cor` to calculate the critical value for a correlation in a prospective framework.
+In the next example we will show the use of the package's function `critical_cor` to calculate the critical values for a correlation in a prospective framework.
 
@@
 The direction of the hypothesis and the test to apply, either $t$-test or $z$-test, should be specified.
-The output will return the critical correlation value, the degrees of freedom and the type of test used.
+The output will return the critical correlation value(s), the degrees of freedom and the type of test used.
 

### 2024-08-20 paper/paper.Rmd
m

@@
       - "Writing - Review & Editing"
+      - "Methodology"
       - "Software"
@@
 In this section, we provide guidance and formulas for computing critical values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's $d$), correlations (Pearson's r), and raw and standardized coefficients in linear models.
-These formulas have been incorporated into R functions of the package "criticalvalue" accessible at: https://github.com/filippogambarota/criticalvalue, and are elaborated upon in the subsequent section.
+These formulas have been incorporated into R functions of the package "criticalESvalue" accessible at: https://github.com/psicostat/criticalESvalue, and are elaborated upon in the subsequent section.
 
@@
 
-In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package "criticalvalue" in R.
-The complete package can be accessed at: https://github.com/filippogambarota/criticalvalue.
+In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package "criticalESvalue" in R.
+The complete package can be accessed at: https://github.com/psicostat/criticalESvalue.
 Here, we demonstrate its application through two examples: one example of a t-test on real data and a computation of the critical effect size values for a correlation from sample size.

### 2024-08-20 paper/paper.tex
m

@@
 The authors made the following contributions.
-Ambra Perugini: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Filippo Gambarota: Writing -- Original Draft Preparation, Writing - Review \& Editing, Methodology; Enrico Toffalini: Writing -- Original Draft Preparation, Writing - Review \& Editing, Supervision; Daniël Lakens: Conceptualization, Writing - Review \& Editing, Supervision; Massimiliano Pastore: Writing - Review \& Editing, Supervision; Livio Finos: Writing - Review \& Editing, Supervision; Psicostat: Writing - Review \& Editing, Supervision; Gianmarco Altoè: Conceptualization, Writing - Review \& Editing, Supervision.
+Ambra Perugini: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing, Methodology, Software; Filippo Gambarota: Writing -- Original Draft Preparation, Writing - Review \& Editing, Methodology, Software; Enrico Toffalini: Writing -- Original Draft Preparation, Writing - Review \& Editing, Software, Supervision; Daniël Lakens: Conceptualization, Writing - Review \& Editing, Supervision; Massimiliano Pastore: Conceptualization, Writing - Review \& Editing, Supervision; Livio Finos: Conceptualization, Writing - Review \& Editing, Supervision; Psicostat: Conceptualization, Writing - Review \& Editing, Supervision; Gianmarco Altoè: Conceptualization, Writing - Review \& Editing, Supervision.
 
@@
 In this section, we provide guidance and formulas for computing critical values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's r), and raw and standardized coefficients in linear models.
-These formulas have been incorporated into R functions of the package ``criticalESvalues'' accessible at: {[}xxxxx{]}, and are elaborated upon in the subsequent section.
+These formulas have been incorporated into R functions of the package ``criticalvalue'' accessible at: \url{https://github.com/filippogambarota/criticalvalue}, and are elaborated upon in the subsequent section.
 
@@
 
-In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package ``criticalESvalues'' in R.
-The complete package can be accessed at: {[}xxxx{]}.
-Here, we demonstrate its application through two examples: one example of a t-test on real data and a computation of the critical value for a correlation from sample size.
-In the Supplementary online materials additional examples can be found for correlation, t-test, paired t-test, linear models and regression coefficients, both from data and from sample size.
+In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package ``criticalvalue'' in R.
+The complete package can be accessed at: \url{https://github.com/filippogambarota/criticalvalue}.
+Here, we demonstrate its application through two examples: one example of a t-test on real data and a computation of the critical effect size values for a correlation from sample size.
+It is important to note that, in general, depending on the researcher's hypotheses, the package allows for the calculation of either two critical effect size values for two-tailed tests or a single critical effect size value for one-tailed tests, for which the direction must be specified.
+In the Supplementary online materials additional examples can be found for correlation, t-test, paired t-test, linear models and regression coefficients.
 
@@
 
-We want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception.
-In this case, it can be easily done with the same procedure using the `t.test' function:
+We want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception. In this case, it can be easily done using the `t.test' function:
 
@@
 \NormalTok{tt }\OtherTok{\textless{}{-}} \FunctionTok{t.test}\NormalTok{(Holz}\SpecialCharTok{$}\NormalTok{t01\_visperc[Holz}\SpecialCharTok{$}\NormalTok{female }\SpecialCharTok{==} \DecValTok{1}\NormalTok{], }
-\NormalTok{             Holz}\SpecialCharTok{$}\NormalTok{t01\_visperc[Holz}\SpecialCharTok{$}\NormalTok{female }\SpecialCharTok{==} \DecValTok{2}\NormalTok{])}
+\NormalTok{             Holz}\SpecialCharTok{$}\NormalTok{t01\_visperc[Holz}\SpecialCharTok{$}\NormalTok{female }\SpecialCharTok{==} \DecValTok{2}\NormalTok{], }\AttributeTok{var.equal =}\NormalTok{ T)}
 \FunctionTok{critical}\NormalTok{(tt)}
@@
 #> 
-#>  Welch Two Sample t-test
+#>  Two Sample t-test
 #> 
 #> data:  Holz$t01_visperc[Holz$female == 1] and Holz$t01_visperc[Holz$female == 2]
-#> t = 1.4095, df = 298.9, p-value = 0.1597
+#> t = 1.4062, df = 299, p-value = 0.1607
 #> alternative hypothesis: true difference in means is not equal to 0
 #> 95 percent confidence interval:
-#>  -0.06419229  0.38823118
+#>  -0.06472157  0.38876046
 #> sample estimates:
@@
 #> |== Effect Size and Critical Value ==| 
-#> d = 0.1623665 dc = ± 0.2269609 bc = ± 0.2262117 
-#> g = 0.1619587 gc = ± 0.2263909
+#> d = 0.1621764 dc = ± 0.2269606 bc = ± 0.226741 
+#> g = 0.1617692 gc = ± 0.2263908
 \end{verbatim}
@@
 The output gives a wide range of values: the Cohen's \(d\) calculated on the data (\(d\)), the critical Cohen's \(d\) (\(dc\)), the numerator of the formula for the critical Cohen's \(d\) (\(bc\)), the Cohen's \(d\) adjusted for small samples (\(g\)) and the critical Cohen's \(d\) adjusted for small samples (\(gc\)).
+The variance is set to be equal, but if that is not the case of your data, you can also run Welch two sample \(t\)-test and obtain the critical effect size values for that.
 
-In the next example we will show the use of the package's function \texttt{critical\_cor} to calculate the critical value for a correlation in a prospective framework.
+In the next example we will show the use of the package's function \texttt{critical\_cor} to calculate the critical values for a correlation in a prospective framework.
 
@@
 The direction of the hypothesis and the test to apply, either \(t\)-test or \(z\)-test, should be specified.
-The output will return the critical correlation value, the degrees of freedom and the type of test used.
+The output will return the critical correlation value(s), the degrees of freedom and the type of test used.
 

### 2024-08-20 paper/paper.Rmd
m

@@
 Reporting critical effect size values may be especially useful when adopting NHST (Null Hypothesis Significance Testing) and the sample size has not been planned a priori, or when there is uncertainty on the expected effect size.
-To assist researchers in calculating critical values, we have developed an R package that allows to report the critical effect size values of model parameters.
+To assist researchers in calculating critical effect size values, we have developed an R package that allows to report the critical effect size values of model parameters.
 The package's functions are implemented for group comparisons, correlations, linear regression, and meta-analysis.
@@
 The "critical effect size values" refers to the smallest statistically significant effect sizes, which depends on the test performed and the characteristics of the sample [@lakens2022sample].
-To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with $n = 20$ participants and uses $p < 0.05$ as the threshold for statistical significance: in this case, the smallest significant effects (critical values) are Pearson's $r = -0.44$ and $r = 0.44$.
-As we will argue, the critical values can serve as a useful complement to aid in the interpretation of study results, communicate their relevance, help assessing their replicability, particularly when statistical significance is relied upon for inference.
+To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with $n = 20$ participants and uses $p < 0.05$ as the threshold for statistical significance: in this case, the smallest significant effects (critical effect size values) are Pearson's $r = -0.44$ and $r = 0.44$.
+As we will argue, the critical effect size values can serve as a useful complement to aid in the interpretation of study results, communicate their relevance, help assessing their replicability, particularly when statistical significance is relied upon for inference.
 This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size or when statistical power may be difficult to determine or deviates from an optimal level.
@@
 Thus, while statistical significance can serve as a potentially useful filter, relying solely on it for inference may introduce risks in interpretation.
-In such cases, we contend that reporting critical values can help put into perspective and convey the relevance of the results.
+In such cases, we contend that reporting critical effect size values can help put into perspective and convey the relevance of the results.
 
-There are two clear benefits of reporting the critical effect size for a corresponding test.
-First, when sample sizes are small, the critical values inform readers about whether the effect sizes that could lead to a rejection of the null hypothesis are in line with realistic expectations.
+There are two clear benefits of reporting the critical effect size value for a corresponding test.
+First, when sample sizes are small, the critical effect size values inform readers about whether the effect sizes that could lead to a rejection of the null hypothesis are in line with realistic expectations.
 If the sample size is small, and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, in these cases, researchers will realized they are not able to collect sufficient data to perform a meaningful test.
-An a-priori power analysis would typically lead to a similar conclusion, but reporting the critical effect sizes focuses the attention more strongly on which effect sizes are reasonable to expect.
-Second, in large studies the critical effect sizes will make clear that trivially small effect sizes will be statistically significant.
+An a-priori power analysis would typically lead to a similar conclusion, but reporting the critical effect size values focuses the attention more strongly on which effect sizes are reasonable to expect.
+Second, in large studies the critical effect size values will make clear that trivially small effect sizes will be statistically significant.
 This will focus the attention of researchers on the difference between statistical significance and practical significance.
@@
 Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to some extent.
-Subsequently, they compute the critical values, revealing Cohen's $d = -0.75$ and $d = 0.75$.
+Subsequently, they compute the critical effect size values, revealing Cohen's $d = -0.75$ and $d = 0.75$.
 This indicates that any estimated |Cohen's $d| < 0.75$ will certainly fail to reach statistical significance.
-By reporting these critical values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance.
+By reporting these critical effect size values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance.
 In essence, this scenario exemplifies a version of the "winner's curse" [@hedges1984estimation], made explicit.
 In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a "winner") to be associated with an overestimated effect size which will probably be deflated in subsequent replications (i.e., the "curse"), especially if such initial finding was obtained on a small sample.
-Knowing from the critical values that a specific result must necessarily be associated with an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.
+Knowing from the critical effect size values that a specific result must necessarily be associated with an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.
 
@@
 It is worth noting that such small effects may potentially reflect just minor artifacts [@wilson2020science], such as subtle experimenter effects or slight non-independence among observations, even in meticulously designed studies.
-While readers may be tempted to interpret any reported effects simply because they attain statistical significance, preliminarily signaling the critical value level can serve as a clear warning that the significance filter might be somehow permissive, thus urging interpretive caution.
+While readers may be tempted to interpret any reported effects simply because they attain statistical significance, preliminarily signaling the critical effect size value level can serve as a clear warning that the significance filter might be somehow permissive, thus urging interpretive caution.
 
@@
 While the focus of meta-analysis is generally on estimating effect sizes with uncertainty, statistical significance is routinely reported and interpreted.
-Thus, in meta-analysis as well, reporting the critical values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.
+Thus, in meta-analysis as well, reporting the critical effect size values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.
 
-To provide more tangible illustrations of the practical application of critical values, let us examine two real-world instances drawn from published research.
+To provide more tangible illustrations of the practical application of critical effect size values, let us examine two real-world instances drawn from published research.
 The first one features a small sample, and has been targeted in the series of replications by the Open Science Collaboration [@open2015estimating], while the second one presents with a very large sample and has already attracted attention in the scientific community.
@@
 
-In this section, we provide guidance and formulas for computing critical values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's $d$), correlations (Pearson's r), and raw and standardized coefficients in linear models.
+In this section, we provide guidance and formulas for computing critical effect size values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's $d$), correlations (Pearson's r), and raw and standardized coefficients in linear models.
 These formulas have been incorporated into R functions of the package "criticalESvalue" accessible at: https://github.com/psicostat/criticalESvalue, and are elaborated upon in the subsequent section.
@@
 Where $t_c$ is the critical $t$ value calculated using a certain $\alpha$ value (e.g., $0.05$) and $n_1 + n_2 - 2$ degrees of freedom.
-Then the $b_c$ is simply divided by $s$ obtaining the standardized critical value.
+Then the $b_c$ is simply divided by $s$ obtaining the standardized critical effect size value.
 For the two-sample case ($s = s_p$ assuming equal variances) Equation \@ref(eq:eq-s-p-twsmp) reported the pooled standard deviation.
@@
 Where $\rho$ is the correlation between the two paired samples.
-Even if the hypothesis testing is computed using $s_D$, we reported the critical value both using the pooled standard deviation and the standard deviation of the differences.
+Even if the hypothesis testing is computed using $s_D$, we reported the critical effect size value both using the pooled standard deviation and the standard deviation of the differences.
 Beyond this difference, the equations are still the same as the previous example.
@@
 Then a $t$ distribution using $n - p - 1$ degrees of freedom where $n$ is the number of observations and $p$ is the number of coefficients beyond the intercept.
-The $t$ value is calculated as reported in Equation \@ref(eq:eq-wald-lm) and the critical value as reported in Equation \@ref(eq:eq-critical-lm).
+The $t$ value is calculated as reported in Equation \@ref(eq:eq-wald-lm) and the critical effect size value as reported in Equation \@ref(eq:eq-critical-lm).
 
@@
 Similarly to standard linear regression, hypothesis testing is performed using Wald $t$ or $z$ tests.
-The calculation of critical values is the same as reported in Equations \@ref(eq:eq-wald-lm) and \@ref(eq:eq-critical-lm).
+The calculation of critical effect size values is the same as reported in Equations \@ref(eq:eq-wald-lm) and \@ref(eq:eq-critical-lm).
 The only difference is that the standard error of the meta-analysis parameters included both the within-studies variances and eventually the between-studies heterogeneity.
@@
 Despite we discussed only linear models, the same approach could be theoretically applied to other types of models such as generalized linear models.
-In fact, we simply need to multiply the critical value of the chosen distribution (e.g., Student $t$ or Standard Normal) by the standard error of the regression coefficient.
+In fact, we simply need to multiply the critical effect size value of the chosen distribution (e.g., Student $t$ or Standard Normal) by the standard error of the regression coefficient.
 
@@
 
-In the next example we will show the use of the package's function `critical_cor` to calculate the critical values for a correlation in a prospective framework.
+In the next example we will show the use of the package's function `critical_cor` to calculate the critical effect size values for a correlation in a prospective framework.
 
@@
 Instead, our proposal serves as a complementary tool, especially beneficial for facilitating the interpretation of results when statistical power deviates from an optimal level (typically falling below, but occasionally exceeding it).
-Interestingly, the critical value can be retrospectively applied even to already published studies.
+Interestingly, the critical effect size value can be retrospectively applied even to already published studies.
 This possible application facilitates potential reframing of the original interpretations.
-Serving as a tool for retrospective analysis, the critical value may enable a reconsideration of the relevance of previously reported findings.
+Serving as a tool for retrospective analysis, the critical effect size value may enable a reconsideration of the relevance of previously reported findings.
 
-An advantage of reporting the critical value is that it can be precisely computed in any scenario, without requiring assumptions about the expected effect size, as is the case with power calculations.
-The critical value represents a directly interpretable benchmark that is especially useful in situations where statistical power is below the desired level and researchers are left otherwise uncertain about how to proceed with the interpretation of a study findings.
+An advantage of reporting the critical effect size value is that it can be precisely computed in any scenario, without requiring assumptions about the expected effect size, as is the case with power calculations.
+The critical effect size value represents a directly interpretable benchmark that is especially useful in situations where statistical power is below the desired level and researchers are left otherwise uncertain about how to proceed with the interpretation of a study findings.
 For example, let us say that we read a published article reporting some effects as statistically significant, while others as not: we suspect that the study may be underpowered, but we are widely uncertain about the magnitude of possible true effects.
 To what extent can the reported results be interpreted, precisely?
-Knowing the critical value provides us with a clear benchmark.
+Knowing the critical effect size value provides us with a clear benchmark.
 Conversely, let us say that an effect achieves significance in a very large sample: researchers tend to draw substantive conclusions based on this.
@@
 If in comparing two groups, such as controls versus treatment, any Cohen's $d > 0.07$ would reach significance, is statistical significance enough to signal a "successful" treatment?
-Maybe yes, even if effects are tiny [e.g., @funder2019evaluating], but knowing the critical value certainly prompts some appropriate interpretive caution.
+Maybe yes, even if effects are tiny [e.g., @funder2019evaluating], but knowing the critical effect size value certainly prompts some appropriate interpretive caution.
 
-Reporting the critical value(s) can also be an efficient way to allow researchers to evaluate which findings are statistically significant.
+Reporting the critical effect size value(s) can also be an efficient way to allow researchers to evaluate which findings are statistically significant.
 For example, in a correlation table researchers customarily add an asterisk to all statistically significant correlations.
@@
 
-Beyond enhancing study design and statistical inferences based on hypothesis tests, reporting critical value(s) can also serve an educational purpose.
+Beyond enhancing study design and statistical inferences based on hypothesis tests, reporting critical effect size value(s) can also serve an educational purpose.
 It underscores how the distinction between a significant and non-significant result is not solely determined by the presence or absence of a true effect, but also by the sample size.
-By highlighting a critical value, researchers can become more aware of the possibility of Type 2 errors when results are non-significant.
-Conversely, in studies with exceedingly large samples and in many meta-analyses, the critical value(s) may serves as a reminder that any observed effect larger than a trivially small value will likely achieve significance.
+By highlighting a critical effect size value, researchers can become more aware of the possibility of Type 2 errors when results are non-significant.
+Conversely, in studies with exceedingly large samples and in many meta-analyses, the critical effect size value(s) may serves as a reminder that any observed effect larger than a trivially small value will likely achieve significance.
 This emphasizes that the mere attainment of statistical significance in a test is not particularly surprising, especially in non-experimental studies.
@@
 Real-case scenarios may not always be that simple.
-Hence, we chose to expand the application of computing critical significance values beyond Cohen's $d$ and correlation to include linear regression with both raw and standardized coefficients.
-This serves as a first step in computing critical values for a wider array of effects encountered in practical scenarios, where linear models and their extensions are commonly utilized for modeling purposes.
+Hence, we chose to expand the application of computing critical effect size values beyond Cohen's $d$ and correlation to include linear regression with both raw and standardized coefficients.
+This serves as a first step in computing critical effect size values for a wider array of effects encountered in practical scenarios, where linear models and their extensions are commonly utilized for modeling purposes.
 A prerequisite is that researchers must be able to identify what parameters in their statistical models reflect the effect sizes of interest, and that they are able to assess their relevance.
@@
 In cases where optimal power can be attained with a sufficiently large sample size for an effect of a specific magnitude of interest, and this is truly determined \textit{a priori}, the interpretation of both significance and non-significance becomes straightforward.
-However, when power analysis did not inform the sample size or when power is likely but undeterminedly low, reporting critical values for the obtained sample can help provide context for interpretation.
-Critical values can be computed and interpreted even retrospectively or for studies that have already been published.
+However, when power analysis did not inform the sample size or when power is likely but undeterminedly low, reporting critical effect size values for the obtained sample can help provide context for interpretation.
+Critical effect size values can be computed and interpreted even retrospectively or for studies that have already been published.
 
-In conclusion, the reporting of the critical value in empirical articles serves as a valuable addition to researchers' toolkit, aimed at augmenting transparency and facilitating the interpretability of their findings.
+In conclusion, the reporting of the critical effect size value in empirical articles serves as a valuable addition to researchers' toolkit, aimed at augmenting transparency and facilitating the interpretability of their findings.
 While not designed to supplant existing practices, it provides a useful aid in interpreting newly presented and previously published results, thus advancing the understanding of research outcomes.

### 2024-08-20 paper/paper.tex
m

@@
 The ``critical effect size values'' refers to the smallest statistically significant effect sizes, which depends on the test performed and the characteristics of the sample (Lakens, 2022).
-To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with \(n = 20\) participants and uses \(p < 0.05\) as the threshold for statistical significance: in this case, the smallest significant effects (critical values) are Pearson's \(r = -0.44\) and \(r = 0.44\).
-As we will argue, the critical values can serve as a useful complement to aid in the interpretation of study results, communicate their relevance, help assessing their replicability, particularly when statistical significance is relied upon for inference.
+To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with \(n = 20\) participants and uses \(p < 0.05\) as the threshold for statistical significance: in this case, the smallest significant effects (critical effect size values) are Pearson's \(r = -0.44\) and \(r = 0.44\).
+As we will argue, the critical effect size values can serve as a useful complement to aid in the interpretation of study results, communicate their relevance, help assessing their replicability, particularly when statistical significance is relied upon for inference.
 This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size or when statistical power may be difficult to determine or deviates from an optimal level.
@@
 Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to some extent.
-Subsequently, they compute the critical values, revealing Cohen's \(d = -0.75\) and \(d = 0.75\).
+Subsequently, they compute the critical effect size values, revealing Cohen's \(d = -0.75\) and \(d = 0.75\).
 This indicates that any estimated \textbar Cohen's \(d| < 0.75\) will certainly fail to reach statistical significance.
-By reporting these critical values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance.
+By reporting these critical effect size values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance.
 In essence, this scenario exemplifies a version of the ``winner's curse'' (Hedges, 1984), made explicit.
 In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a ``winner'') to be associated with an overestimated effect size which will probably be deflated in subsequent replications (i.e., the ``curse''), especially if such initial finding was obtained on a small sample.
-Knowing from the critical values that a specific result must necessarily be associated with an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.
+Knowing from the critical effect size values that a specific result must necessarily be associated with an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.
 
@@
 It is worth noting that such small effects may potentially reflect just minor artifacts (Wilson et al., 2020), such as subtle experimenter effects or slight non-independence among observations, even in meticulously designed studies.
-While readers may be tempted to interpret any reported effects simply because they attain statistical significance, preliminarily signaling the critical value level can serve as a clear warning that the significance filter might be somehow permissive, thus urging interpretive caution.
+While readers may be tempted to interpret any reported effects simply because they attain statistical significance, preliminarily signaling the critical effect size value level can serve as a clear warning that the significance filter might be somehow permissive, thus urging interpretive caution.
 
@@
 While the focus of meta-analysis is generally on estimating effect sizes with uncertainty, statistical significance is routinely reported and interpreted.
-Thus, in meta-analysis as well, reporting the critical values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.
+Thus, in meta-analysis as well, reporting the critical effect size values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.
 
-To provide more tangible illustrations of the practical application of critical values, let us examine two real-world instances drawn from published research.
+To provide more tangible illustrations of the practical application of critical effect size values, let us examine two real-world instances drawn from published research.
 The first one features a small sample, and has been targeted in the series of replications by the Open Science Collaboration (Collaboration, 2015), while the second one presents with a very large sample and has already attracted attention in the scientific community.
@@
 
-In this section, we provide guidance and formulas for computing critical values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's r), and raw and standardized coefficients in linear models.
-These formulas have been incorporated into R functions of the package ``criticalvalue'' accessible at: \url{https://github.com/filippogambarota/criticalvalue}, and are elaborated upon in the subsequent section.
+In this section, we provide guidance and formulas for computing critical effect size values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's r), and raw and standardized coefficients in linear models.
+These formulas have been incorporated into R functions of the package ``criticalESvalue'' accessible at: \url{https://github.com/psicostat/criticalESvalue}, and are elaborated upon in the subsequent section.
 
@@
 Where \(t_c\) is the critical \(t\) value calculated using a certain \(\alpha\) value (e.g., \(0.05\)) and \(n_1 + n_2 - 2\) degrees of freedom.
-Then the \(b_c\) is simply divided by \(s\) obtaining the standardized critical value.
+Then the \(b_c\) is simply divided by \(s\) obtaining the standardized critical effect size value.
 For the two-sample case (\(s = s_p\) assuming equal variances) Equation \eqref{eq:eq-s-p-twsmp} reported the pooled standard deviation.
@@
 Where \(\rho\) is the correlation between the two paired samples.
-Even if the hypothesis testing is computed using \(s_D\), we reported the critical value both using the pooled standard deviation and the standard deviation of the differences.
+Even if the hypothesis testing is computed using \(s_D\), we reported the critical effect size value both using the pooled standard deviation and the standard deviation of the differences.
 Beyond this difference, the equations are still the same as the previous example.
@@
 Then a \(t\) distribution using \(n - p - 1\) degrees of freedom where \(n\) is the number of observations and \(p\) is the number of coefficients beyond the intercept.
-The \(t\) value is calculated as reported in Equation \eqref{eq:eq-wald-lm} and the critical value as reported in Equation \eqref{eq:eq-critical-lm}.
+The \(t\) value is calculated as reported in Equation \eqref{eq:eq-wald-lm} and the critical effect size value as reported in Equation \eqref{eq:eq-critical-lm}.
 
@@
 Similarly to standard linear regression, hypothesis testing is performed using Wald \(t\) or \(z\) tests.
-The calculation of critical values is the same as reported in Equations \eqref{eq:eq-wald-lm} and \eqref{eq:eq-critical-lm}.
+The calculation of critical effect size values is the same as reported in Equations \eqref{eq:eq-wald-lm} and \eqref{eq:eq-critical-lm}.
 The only difference is that the standard error of the meta-analysis parameters included both the within-studies variances and eventually the between-studies heterogeneity.
@@
 Despite we discussed only linear models, the same approach could be theoretically applied to other types of models such as generalized linear models.
-In fact, we simply need to multiply the critical value of the chosen distribution (e.g., Student \(t\) or Standard Normal) by the standard error of the regression coefficient.
+In fact, we simply need to multiply the critical effect size value of the chosen distribution (e.g., Student \(t\) or Standard Normal) by the standard error of the regression coefficient.
 
@@
 
-In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package ``criticalvalue'' in R.
-The complete package can be accessed at: \url{https://github.com/filippogambarota/criticalvalue}.
+In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package ``criticalESvalue'' in R.
+The complete package can be accessed at: \url{https://github.com/psicostat/criticalESvalue}.
 Here, we demonstrate its application through two examples: one example of a t-test on real data and a computation of the critical effect size values for a correlation from sample size.
@@
 
-In the next example we will show the use of the package's function \texttt{critical\_cor} to calculate the critical values for a correlation in a prospective framework.
+In the next example we will show the use of the package's function \texttt{critical\_cor} to calculate the critical effect size values for a correlation in a prospective framework.
 
@@
 If in comparing two groups, such as controls versus treatment, any Cohen's \(d > 0.07\) would reach significance, is statistical significance enough to signal a ``successful'' treatment?
-Maybe yes, even if effects are tiny (e.g., Funder \& Ozer, 2019), but knowing the critical value certainly prompts some appropriate interpretive caution.
+Maybe yes, even if effects are tiny (e.g., Funder \& Ozer, 2019), but knowing the critical effect size value certainly prompts some appropriate interpretive caution.
 
-Reporting the critical value(s) can also be an efficient way to allow researchers to evaluate which findings are statistically significant.
+Reporting the critical effect size value(s) can also be an efficient way to allow researchers to evaluate which findings are statistically significant.
 For example, in a correlation table researchers customarily add an asterisk to all statistically significant correlations.
@@
 Real-case scenarios may not always be that simple.
-Hence, we chose to expand the application of computing critical significance values beyond Cohen's \(d\) and correlation to include linear regression with both raw and standardized coefficients.
-This serves as a first step in computing critical values for a wider array of effects encountered in practical scenarios, where linear models and their extensions are commonly utilized for modeling purposes.
+Hence, we chose to expand the application of computing critical effect size values beyond Cohen's \(d\) and correlation to include linear regression with both raw and standardized coefficients.
+This serves as a first step in computing critical effect size values for a wider array of effects encountered in practical scenarios, where linear models and their extensions are commonly utilized for modeling purposes.
 A prerequisite is that researchers must be able to identify what parameters in their statistical models reflect the effect sizes of interest, and that they are able to assess their relevance.

### 2024-08-20 paper/paper.Rmd
m

@@
 These values are highly applicable in various scenarios, as they do not require specifying a plausible effect size of interest.
-Reporting critical effect size values may be especially useful when adopting NHST (Null Hypothesis Significance Testing) and the sample size has not been planned a priori, or when there is uncertainty on the expected effect size.
+Reporting critical effect size values may be especially useful when adopting NHST (Null Hypothesis Significance Testing) and the sample size has not been planned a priori, or when there is uncertainty about the expected effect size.
 To assist researchers in calculating critical effect size values, we have developed an R package that allows to report the critical effect size values of model parameters.
@@
 keywords          : critical effect size values, statistical significance, fucntions implementation 
-wordcount         : "X"
+wordcount         : "3900"
 
@@
 
-In the present paper, we suggest that researchers systematically report the critical effect size values when they make inference using statistical significance.
-The "critical effect size values" refers to the smallest statistically significant effect sizes, which depends on the test performed and the characteristics of the sample [@lakens2022sample].
+In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference using statistical significance.
+The "critical effect size values" refers to the smallest statistically significant effect sizes, which depend on the test performed and the characteristics of the sample [@lakens2022sample].
 To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with $n = 20$ participants and uses $p < 0.05$ as the threshold for statistical significance: in this case, the smallest significant effects (critical effect size values) are Pearson's $r = -0.44$ and $r = 0.44$.
-As we will argue, the critical effect size values can serve as a useful complement to aid in the interpretation of study results, communicate their relevance, help assessing their replicability, particularly when statistical significance is relied upon for inference.
-This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size or when statistical power may be difficult to determine or deviates from an optimal level.
+As we will argue, critical effect size values can serve as a useful complement to aid in the interpretation of study results, to communicate their relevance, and to help assess their replicability, particularly when statistical significance is relied upon for inference.
+This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size, or when statistical power may be difficult to determine or deviates from an optimal level.
 
-Despite longstanding criticisms, the Null Hypothesis Significance Testing (NHST) remains a prominent approach for statistical inference in science [@cohen1994; @gigerenzer2004null; @krueger2001null].
+Despite longstanding criticisms, Null Hypothesis Significance Testing (NHST) remains a prominent approach for statistical inference in science [@cohen1994; @gigerenzer2004null; @krueger2001null].
 Over the past decades, numerous proposals have emerged to enhance inference based on this approach.
-These include conducting power analyses to determine sample size ($n$) a priori based on theory-informed target effects, and reporting of effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance [@transue2019apa].
+These include conducting power analyses to determine sample size ($n$) a priori based on theory-informed target effects, and reporting effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance [@transue2019apa].
 Even for the most skeptical, there are suggestions that NHST may possess merits of its own.
-For instance, @wilson2020science propose that it could serve as a useful preliminary filter utilized by "original science" for screening potentially interesting effects, that should subsequently be validated by "replication science".
+For instance, @wilson2020science propose that it could serve as a useful preliminary filter utilized by "original science" for screening potentially interesting effects, which should subsequently be validated by "replication science".
 
-In psychology, power is often severely limited for medium to small effect sizes [@szucs2017empirical], which are generally expected to represent true replicable findings.
+In psychology, statistical power is often severely limited for medium to small effect sizes [@szucs2017empirical], which are generally expected to represent true replicable findings.
 In fact, resource constraints have been authoritatively listed among acceptable justifications for sample size  [@lakens2022sample], and this scenario is arguably widespread in psychology.
@@
 On the other hand, very large samples are occasionally available in psychological science, albeit less frequently.
-However, as some have argued, excessively large sample sizes risk creating an "everything-is-significant" scenario, where researchers report and discuss practically negligible effects solely because they are statistically significant, regardless of theoretical relevance or whether they just reflect minor procedural artifacts [@wilson2020science].
+However, as some have argued, excessively large sample sizes risk creating an "everything-is-significant" scenario, where researchers report and discuss practically negligible effects solely because they are statistically significant, regardless of theoretical relevance or whether they merely reflect minor procedural artifacts [@wilson2020science].
 Thus, while statistical significance can serve as a potentially useful filter, relying solely on it for inference may introduce risks in interpretation.
@@
 
-There are two clear benefits of reporting the critical effect size value for a corresponding test.
-First, when sample sizes are small, the critical effect size values inform readers about whether the effect sizes that could lead to a rejection of the null hypothesis are in line with realistic expectations.
-If the sample size is small, and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, in these cases, researchers will realized they are not able to collect sufficient data to perform a meaningful test.
+There are two clear benefits to reporting the critical effect size value for a corresponding test.
+First, when sample sizes are small, the critical effect size values inform readers about whether the effect sizes that could lead to rejecting the null hypothesis are in line with realistic expectations.
+If the sample size is small and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, in these cases, researchers will realize they are not able to collect sufficient data to perform a meaningful test.
 An a-priori power analysis would typically lead to a similar conclusion, but reporting the critical effect size values focuses the attention more strongly on which effect sizes are reasonable to expect.
-Second, in large studies the critical effect size values will make clear that trivially small effect sizes will be statistically significant.
+Second, in large studies the critical effect size values will make it clear that trivially small effect sizes will be statistically significant.
 This will focus the attention of researchers on the difference between statistical significance and practical significance.
@@
 In essence, this scenario exemplifies a version of the "winner's curse" [@hedges1984estimation], made explicit.
-In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a "winner") to be associated with an overestimated effect size which will probably be deflated in subsequent replications (i.e., the "curse"), especially if such initial finding was obtained on a small sample.
-Knowing from the critical effect size values that a specific result must necessarily be associated with an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.
+In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a "winner") to be associated with an overestimated effect size which will likely be deflated in subsequent replications (i.e., the "curse"), especially if such initial finding was obtained with a small sample.
+Knowing from the critical effect size values that a specific result must necessarily be associated with either an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.
 
 In a second scenario, imagine researchers who gain access to a very large archival dataset ($n = 5,000$) and decide to explore some bivariate correlations.
-With such a large sample size, researchers are not constrained by sample size limitations and could instead just focus on effect sizes.
-However, they opt to legitimately use statistical significance of a two-tailed test as their primary criterion for whether to comment on effects in their further discussions.
+With such a large sample size, researchers are not constrained by sample size limitations and could instead focus on effect sizes.
+However, they legitimately  opt to use statistical significance of a two-tailed test as their primary criterion for whether to comment on effects in their further discussions.
 In this context, the critical effect size values for significance are determined to be correlations of $r = \pm 0.03$.
@@
 It is worth noting that such small effects may potentially reflect just minor artifacts [@wilson2020science], such as subtle experimenter effects or slight non-independence among observations, even in meticulously designed studies.
-While readers may be tempted to interpret any reported effects simply because they attain statistical significance, preliminarily signaling the critical effect size value level can serve as a clear warning that the significance filter might be somehow permissive, thus urging interpretive caution.
+While readers may be tempted to interpret any reported effects simply because they attain statistical significance, signaling the critical effect size value level beforehand can serve as a clear warning that the significance filter might be overly permissive, thus urging interpretive caution.
 
@@
 While the focus of meta-analysis is generally on estimating effect sizes with uncertainty, statistical significance is routinely reported and interpreted.
-Thus, in meta-analysis as well, reporting the critical effect size values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.
+Thus, in meta-analysis, reporting the critical effect size values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.
 
-To provide more tangible illustrations of the practical application of critical effect size values, let us examine two real-world instances drawn from published research.
+To provide more tangible illustrations of the practical applications of critical effect size values, let us examine two real-world instances drawn from published research.
 The first one features a small sample, and has been targeted in the series of replications by the Open Science Collaboration [@open2015estimating], while the second one presents with a very large sample and has already attracted attention in the scientific community.

### 2024-08-20 paper/paper.tex
m

@@
 {}{\PatchFailed}
-\keywords{critical effect size values, statistical significance, fucntions implementation\newline\indent Word count: X}
+\keywords{critical effect size values, statistical significance, fucntions implementation\newline\indent Word count: 3900}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 
-In the present paper, we suggest that researchers systematically report the critical effect size values when they make inference using statistical significance.
-The ``critical effect size values'' refers to the smallest statistically significant effect sizes, which depends on the test performed and the characteristics of the sample (Lakens, 2022).
+In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference using statistical significance.
+The ``critical effect size values'' refers to the smallest statistically significant effect sizes, which depend on the test performed and the characteristics of the sample (Lakens, 2022).
 To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with \(n = 20\) participants and uses \(p < 0.05\) as the threshold for statistical significance: in this case, the smallest significant effects (critical effect size values) are Pearson's \(r = -0.44\) and \(r = 0.44\).
-As we will argue, the critical effect size values can serve as a useful complement to aid in the interpretation of study results, communicate their relevance, help assessing their replicability, particularly when statistical significance is relied upon for inference.
-This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size or when statistical power may be difficult to determine or deviates from an optimal level.
+As we will argue, critical effect size values can serve as a useful complement to aid in the interpretation of study results, to communicate their relevance, and to help assess their replicability, particularly when statistical significance is relied upon for inference.
+This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size, or when statistical power may be difficult to determine or deviates from an optimal level.
 
-Despite longstanding criticisms, the Null Hypothesis Significance Testing (NHST) remains a prominent approach for statistical inference in science (Cohen, 1994; Gigerenzer et al., 2004; Krueger, 2001).
+Despite longstanding criticisms, Null Hypothesis Significance Testing (NHST) remains a prominent approach for statistical inference in science (Cohen, 1994; Gigerenzer et al., 2004; Krueger, 2001).
 Over the past decades, numerous proposals have emerged to enhance inference based on this approach.
-These include conducting power analyses to determine sample size (\(n\)) a priori based on theory-informed target effects, and reporting of effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance (Transue, 2019).
+These include conducting power analyses to determine sample size (\(n\)) a priori based on theory-informed target effects, and reporting effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance (Transue, 2019).
 Even for the most skeptical, there are suggestions that NHST may possess merits of its own.
 For instance, Wilson et al.
-(2020) propose that it could serve as a useful preliminary filter utilized by ``original science'' for screening potentially interesting effects, that should subsequently be validated by ``replication science''.
+(2020) propose that it could serve as a useful preliminary filter utilized by ``original science'' for screening potentially interesting effects, which should subsequently be validated by ``replication science''.
 
-In psychology, power is often severely limited for medium to small effect sizes (Szucs \& Ioannidis, 2017), which are generally expected to represent true replicable findings.
+In psychology, statistical power is often severely limited for medium to small effect sizes (Szucs \& Ioannidis, 2017), which are generally expected to represent true replicable findings.
 In fact, resource constraints have been authoritatively listed among acceptable justifications for sample size (Lakens, 2022), and this scenario is arguably widespread in psychology.
@@
 On the other hand, very large samples are occasionally available in psychological science, albeit less frequently.
-However, as some have argued, excessively large sample sizes risk creating an ``everything-is-significant'' scenario, where researchers report and discuss practically negligible effects solely because they are statistically significant, regardless of theoretical relevance or whether they just reflect minor procedural artifacts (Wilson et al., 2020).
+However, as some have argued, excessively large sample sizes risk creating an ``everything-is-significant'' scenario, where researchers report and discuss practically negligible effects solely because they are statistically significant, regardless of theoretical relevance or whether they merely reflect minor procedural artifacts (Wilson et al., 2020).
 Thus, while statistical significance can serve as a potentially useful filter, relying solely on it for inference may introduce risks in interpretation.
@@
 In essence, this scenario exemplifies a version of the ``winner's curse'' (Hedges, 1984), made explicit.
-In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a ``winner'') to be associated with an overestimated effect size which will probably be deflated in subsequent replications (i.e., the ``curse''), especially if such initial finding was obtained on a small sample.
-Knowing from the critical effect size values that a specific result must necessarily be associated with an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.
+In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a ``winner'') to be associated with an overestimated effect size which will likely be deflated in subsequent replications (i.e., the ``curse''), especially if such initial finding was obtained with a small sample.
+Knowing from the critical effect size values that a specific result must necessarily be associated with either an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.
 
 In a second scenario, imagine researchers who gain access to a very large archival dataset (\(n = 5,000\)) and decide to explore some bivariate correlations.
-With such a large sample size, researchers are not constrained by sample size limitations and could instead just focus on effect sizes.
-However, they opt to legitimately use statistical significance of a two-tailed test as their primary criterion for whether to comment on effects in their further discussions.
+With such a large sample size, researchers are not constrained by sample size limitations and could instead focus on effect sizes.
+However, they legitimately opt to use statistical significance of a two-tailed test as their primary criterion for whether to comment on effects in their further discussions.
 In this context, the critical effect size values for significance are determined to be correlations of \(r = \pm 0.03\).
@@
 It is worth noting that such small effects may potentially reflect just minor artifacts (Wilson et al., 2020), such as subtle experimenter effects or slight non-independence among observations, even in meticulously designed studies.
-While readers may be tempted to interpret any reported effects simply because they attain statistical significance, preliminarily signaling the critical effect size value level can serve as a clear warning that the significance filter might be somehow permissive, thus urging interpretive caution.
+While readers may be tempted to interpret any reported effects simply because they attain statistical significance, signaling the critical effect size value level beforehand can serve as a clear warning that the significance filter might be overly permissive, thus urging interpretive caution.
 
@@
 While the focus of meta-analysis is generally on estimating effect sizes with uncertainty, statistical significance is routinely reported and interpreted.
-Thus, in meta-analysis as well, reporting the critical effect size values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.
+Thus, in meta-analysis, reporting the critical effect size values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.
 
-To provide more tangible illustrations of the practical application of critical effect size values, let us examine two real-world instances drawn from published research.
+To provide more tangible illustrations of the practical applications of critical effect size values, let us examine two real-world instances drawn from published research.
 The first one features a small sample, and has been targeted in the series of replications by the Open Science Collaboration (Collaboration, 2015), while the second one presents with a very large sample and has already attracted attention in the scientific community.

### 2024-08-20 paper/paper.Rmd
m

@@
 Over the past decades, numerous proposals have emerged to enhance inference based on this approach.
-These include conducting power analyses to determine sample size ($n$) a priori based on theory-informed target effects, and reporting effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance [@transue2019apa].
+These include conducting power analyses to determine sample size ($n$) \textit{a priori} based on theory-informed target effects, and reporting effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance [@transue2019apa].
 Even for the most skeptical, there are suggestions that NHST may possess merits of its own.
@@
 To provide more tangible illustrations of the practical applications of critical effect size values, let us examine two real-world instances drawn from published research.
-The first one features a small sample, and has been targeted in the series of replications by the Open Science Collaboration [@open2015estimating], while the second one presents with a very large sample and has already attracted attention in the scientific community.
+The first one features a small sample and has been targeted in the series of replications by the Open Science Collaboration [@open2015estimating], while the second one presents with a very large sample and has already attracted attention in the scientific community.
 The first case pertains to "Study 5" conducted by @mccrea2008self.
@@
 With a modest sample size of 28 participants split into two groups (13 and 15 participants, respectively), a directional one-tailed t-test was conducted on the percentage of correct answers to compare the groups.
-The resulting effect was statistically significant, $t(26) = 1.87, p (one-tailed) < .05$.
+The resulting effect was statistically significant, $t(26) = 1.87$, $p$ (one-tailed) $< .05$.
 Based on the reported data, Cohen's $d = 0.736$, indicating a considerable effect size.
@@
 This is not intended to replace other strategies aimed at enhancing the NHST approach to inference.
-Such strategies, such as the emphasis on estimating effect sizes with confidence intervals [@transue2019apa] or the a priori planning for statistical power are valuable in their own right.
+Such strategies, such as the emphasis on estimating effect sizes with confidence intervals [@transue2019apa] or the \textit{a priori} planning for statistical power are valuable in their own right.
 Instead, our proposal serves as a complementary tool, especially beneficial for facilitating the interpretation of results when statistical power deviates from an optimal level (typically falling below, but occasionally exceeding it).
@@
 This possible application facilitates potential reframing of the original interpretations.
-Serving as a tool for retrospective analysis, the critical effect size value may enable a reconsideration of the relevance of previously reported findings.
+Serving as a tool for retrospective analysis, critical effect size values may enable a reconsideration of the relevance of previously reported findings.
 
@@
 By highlighting a critical effect size value, researchers can become more aware of the possibility of Type 2 errors when results are non-significant.
-Conversely, in studies with exceedingly large samples and in many meta-analyses, the critical effect size value(s) may serves as a reminder that any observed effect larger than a trivially small value will likely achieve significance.
+Conversely, in studies with exceedingly large samples and in many meta-analyses, the critical effect size value(s) may serve as a reminder that any observed effect larger than a trivially small value will likely achieve significance.
 This emphasizes that the mere attainment of statistical significance in a test is not particularly surprising, especially in non-experimental studies.
@@
 
-In conclusion, the reporting of the critical effect size value in empirical articles serves as a valuable addition to researchers' toolkit, aimed at augmenting transparency and facilitating the interpretability of their findings.
+In conclusion, reporting critical effect size values in empirical articles serves as a valuable addition to researchers' toolkit, aimed at augmenting transparency and facilitating the interpretability of their findings.
 While not designed to supplant existing practices, it provides a useful aid in interpreting newly presented and previously published results, thus advancing the understanding of research outcomes.

### 2024-08-20 paper/paper.tex
m

@@
 Over the past decades, numerous proposals have emerged to enhance inference based on this approach.
-These include conducting power analyses to determine sample size (\(n\)) a priori based on theory-informed target effects, and reporting effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance (Transue, 2019).
+These include conducting power analyses to determine sample size (\(n\)) \textit{a priori} based on theory-informed target effects, and reporting effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance (Transue, 2019).
 Even for the most skeptical, there are suggestions that NHST may possess merits of its own.
@@
 To provide more tangible illustrations of the practical applications of critical effect size values, let us examine two real-world instances drawn from published research.
-The first one features a small sample, and has been targeted in the series of replications by the Open Science Collaboration (Collaboration, 2015), while the second one presents with a very large sample and has already attracted attention in the scientific community.
+The first one features a small sample and has been targeted in the series of replications by the Open Science Collaboration (Collaboration, 2015), while the second one presents with a very large sample and has already attracted attention in the scientific community.
 The first case pertains to ``Study 5'' conducted by McCrea (2008).
@@
 With a modest sample size of 28 participants split into two groups (13 and 15 participants, respectively), a directional one-tailed t-test was conducted on the percentage of correct answers to compare the groups.
-The resulting effect was statistically significant, \(t(26) = 1.87, p (one-tailed) < .05\).
+The resulting effect was statistically significant, \(t(26) = 1.87\), \(p\) (one-tailed) \(< .05\).
 Based on the reported data, Cohen's \(d = 0.736\), indicating a considerable effect size.
@@
 This is not intended to replace other strategies aimed at enhancing the NHST approach to inference.
-Such strategies, such as the emphasis on estimating effect sizes with confidence intervals (Transue, 2019) or the a priori planning for statistical power are valuable in their own right.
+Such strategies, such as the emphasis on estimating effect sizes with confidence intervals (Transue, 2019) or the \textit{a priori} planning for statistical power are valuable in their own right.
 Instead, our proposal serves as a complementary tool, especially beneficial for facilitating the interpretation of results when statistical power deviates from an optimal level (typically falling below, but occasionally exceeding it).

### 2024-08-21 paper/paper.Rmd
m

@@
 In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference using statistical significance.
-The "critical effect size values" refers to the smallest statistically significant effect sizes, which depend on the test performed and the characteristics of the sample [@lakens2022sample].
+The "critical effect size value(s)" refers to the smallest statistically significant effect sizes, which depend on the test performed and the characteristics of the sample [@lakens2022sample].
 To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with $n = 20$ participants and uses $p < 0.05$ as the threshold for statistical significance: in this case, the smallest significant effects (critical effect size values) are Pearson's $r = -0.44$ and $r = 0.44$.
@@
 Over the past decades, numerous proposals have emerged to enhance inference based on this approach.
-These include conducting power analyses to determine sample size ($n$) \textit{a priori} based on theory-informed target effects, and reporting effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance [@transue2019apa].
+These include conducting power analyses to determine sample size \textit{a priori} based on theory-informed target effects, and reporting effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance [@transue2019apa].
 Even for the most skeptical, there are suggestions that NHST may possess merits of its own.
@@
 
-There are two clear benefits to reporting the critical effect size value for a corresponding test.
+There are two clear benefits to reporting the critical effect size value(s) for a corresponding test.
 First, when sample sizes are small, the critical effect size values inform readers about whether the effect sizes that could lead to rejecting the null hypothesis are in line with realistic expectations.
 If the sample size is small and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, in these cases, researchers will realize they are not able to collect sufficient data to perform a meaningful test.
-An a-priori power analysis would typically lead to a similar conclusion, but reporting the critical effect size values focuses the attention more strongly on which effect sizes are reasonable to expect.
+An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values focuses the attention more strongly on which effect sizes are reasonable to expect.
 Second, in large studies the critical effect size values will make it clear that trivially small effect sizes will be statistically significant.
@@
 Instead, our proposal serves as a complementary tool, especially beneficial for facilitating the interpretation of results when statistical power deviates from an optimal level (typically falling below, but occasionally exceeding it).
-Interestingly, the critical effect size value can be retrospectively applied even to already published studies.
+Interestingly, critical effect size values can be retrospectively applied even to already published studies.
 This possible application facilitates potential reframing of the original interpretations.
@@
 
-An advantage of reporting the critical effect size value is that it can be precisely computed in any scenario, without requiring assumptions about the expected effect size, as is the case with power calculations.
+An advantage of reporting critical effect size values is that they can be precisely computed in any scenario, without requiring assumptions about the expected effect size, as is the case with power calculations.
 The critical effect size value represents a directly interpretable benchmark that is especially useful in situations where statistical power is below the desired level and researchers are left otherwise uncertain about how to proceed with the interpretation of a study findings.
@@
 
-Beyond enhancing study design and statistical inferences based on hypothesis tests, reporting critical effect size value(s) can also serve an educational purpose.
+Beyond enhancing study design and statistical inferences based on hypothesis tests, reporting critical effect size values can also serve an educational purpose.
 It underscores how the distinction between a significant and non-significant result is not solely determined by the presence or absence of a true effect, but also by the sample size.

### 2024-08-21 paper/paper.tex
m

@@
 In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference using statistical significance.
-The ``critical effect size values'' refers to the smallest statistically significant effect sizes, which depend on the test performed and the characteristics of the sample (Lakens, 2022).
+The ``critical effect size value(s)'' refers to the smallest statistically significant effect sizes, which depend on the test performed and the characteristics of the sample (Lakens, 2022).
 To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with \(n = 20\) participants and uses \(p < 0.05\) as the threshold for statistical significance: in this case, the smallest significant effects (critical effect size values) are Pearson's \(r = -0.44\) and \(r = 0.44\).
@@
 Over the past decades, numerous proposals have emerged to enhance inference based on this approach.
-These include conducting power analyses to determine sample size (\(n\)) \textit{a priori} based on theory-informed target effects, and reporting effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance (Transue, 2019).
+These include conducting power analyses to determine sample size \textit{a priori} based on theory-informed target effects, and reporting effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance (Transue, 2019).
 Even for the most skeptical, there are suggestions that NHST may possess merits of its own.
